---
title: "Periodic and conditional survival trends in prostate, testicular and penile cancers in the nordic countries marking timing of improvements"
subtitle: "Full R code"
author: "Filip Tichanek, Asta Försti, Akseli Hemminki, Otto Hemminki, Kari Hemminki1"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document: 
    toc: yes
    toc_float: yes
    theme: journal
    fig_caption: yes
    keep_md: yes
    number_sections: yes
  word_document: default
  toc: yes
  toc_depth: 5
  number_sections: yes
  pdf_document: default
urlcolor: blue
---

\tableofcontents 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Please, **cite as** [original article](*TO BE ADDED*) published in [*TO BE ADDED*]

Full citation:

[*TO BE ADDED*]

# Introduction

This report shows source data, their wrangling, fitting statistical models, outputs of the models, and visualizations of both data and probabilistic models. 

All analyses were done using [R sofware](https://www.r-project.org/) in [RStudio](https://www.rstudio.com/) interface.

## Software and Packages

Software versions.

Uploading R packages and their version

```{r message=FALSE, warning=FALSE,error=F}
rm(list=ls()) # removing all data and objects from R environment
library(stringr)
library(brms)
library(dplyr)
library(sessioninfo)
library(cowplot)
```




# Incidence and mortality

```{r}
# setting parameters for smoothing
spar=0.3
knot=12

# colors ---------------------------------------------------------

cola<-c(
  rgb(1,0.1,0.1,alpha=1),
  rgb(0.1,0.1,1,alpha=1),
  rgb(0, 0.6, 0.3,alpha=1),
  rgb(0.7,0.7,0.1,alpha=1))

colb<-c(
  rgb(1,0.1,0.1,alpha=0.2),
  rgb(0.1,0.1,1,alpha=0.2),
  rgb(0, 0.6, 0.3,alpha=0.2),
  rgb(0.7,0.7,0.1,alpha=0.2))

colc<-c(
  rgb(1,0.1,0.1,alpha=0.8),
  rgb(0.1,0.1,1,alpha=0.8),
  rgb(0, 0.6, 0.3,alpha=0.8),
  rgb(0.7,0.7,0.1,alpha=0.8))

```

## Data upload 

Data were downloaded from [Nordcan](https://nordcan.iarc.fr/en/)


```{r}
## testes data ----------------------------------------------------------
urlfile="https://raw.githubusercontent.com/filip-tichanek/nord_male/main/source-data/testes_inc_mor.csv"
testes_inc_mor<-read.csv(url(urlfile),sep=",")

colnam<-testes_inc_mor[,1]
testes_inc_mor<-data.frame(t(testes_inc_mor))[-1,]
colnames(testes_inc_mor) <- colnam
testes_inc_mor$year <- 1943:2020
### Subset-ting years 1961-2020
testes_inc_mor<-testes_inc_mor[testes_inc_mor$year>1960,]
### Removing space character and converting characters to numbers
x=1;repeat{
testes_inc_mor[,x]<-str_trim(testes_inc_mor[,x])
testes_inc_mor[,x]<-as.numeric(testes_inc_mor[,x])
x=x+1;if(x>9){break}}
summary(testes_inc_mor)

## prostate data ----------------------------------------------------------
urlfile="https://raw.githubusercontent.com/filip-tichanek/nord_male/main/source-data/prostate_inc_mor.csv"
prostate_inc_mor<-read.csv(url(urlfile),sep=",")
colnam<-prostate_inc_mor[,1]
prostate_inc_mor<-data.frame(t(prostate_inc_mor))[-1,]
colnames(prostate_inc_mor) <- colnam
prostate_inc_mor$year <- 1943:2020
### Subset-ting years 1961-2020
prostate_inc_mor<-prostate_inc_mor[prostate_inc_mor$year>1960,]
### Removing space character and converting characters to numbers
x=1;repeat{
  prostate_inc_mor[,x]<-str_trim(prostate_inc_mor[,x])
  prostate_inc_mor[,x]<-as.numeric(prostate_inc_mor[,x])
  x=x+1;if(x>9){break}}
summary(prostate_inc_mor)

## penis data ----------------------------------------------------------
urlfile="https://raw.githubusercontent.com/filip-tichanek/nord_male/main/source-data/penis_inc_mor.csv"
penis_inc_mor<-read.csv(url(urlfile),sep=",")
colnam<-penis_inc_mor[,1]
penis_inc_mor<-data.frame(t(penis_inc_mor))[-1,]
colnames(penis_inc_mor) <- colnam
penis_inc_mor$year <- 1943:2020
### Subseting years 1961-2020
penis_inc_mor<-penis_inc_mor[penis_inc_mor$year>1960,]
### Removing space character and converting characters to numbers
x=1;repeat{
  penis_inc_mor[,x]<-str_trim(penis_inc_mor[,x])
  penis_inc_mor[,x]<-as.numeric(penis_inc_mor[,x])
  x=x+1;if(x>9){break}}
summary(penis_inc_mor)
```


## Suppl. Figure - Incidence and Moratlity

```{r, fig.height=6.5,fig.width=8.57, fig.cap="Incidence and Mortality in male cancers"}
# Plotting - incidence -------------------------
## General setting and titles of plots
m <- matrix(c(1 ,2 ,3,
              4 ,5 ,6,
              7, 8, 9), nrow = 3, ncol =3 ,byrow = TRUE)
layout(mat = m,heights = c(0.03,0.485,0.485))
par(mgp=c(1.6,0.8,0))
par(mar=c(0,0,0,0))

plot(NULL, axes=FALSE,xlab="",ylab="",xlim=c(-1,1),ylim=c(-0.85,0.85))
text(0.05,-0.,"Testis",cex=1.8,font=3,xpd=TRUE)

plot(NULL, axes=FALSE,xlab="",ylab="",xlim=c(-1,1),ylim=c(-0.85,0.85))
text(0.05,-0.,"Prostate",cex=1.8,font=3,xpd=TRUE)

plot(NULL, axes=FALSE,xlab="",ylab="",xlim=c(-1,1),ylim=c(-0.85,0.85))
text(0.05,-0.,"Penis",cex=1.8,font=3,xpd=TRUE)



## Plot of testes cancer ----------------
par(mgp=c(0.1,0.7,0))
par(mar=c(2,2,0,0))

range<-c(0,12);scal<-range[2]-range[1]
xrange<-c(1961,2020)

plot(NULL,xlim=xrange,ylim=c(range[1],range[2]),xlab="",ylab=""
     ,las=1, axes=FALSE)
rect(xrange[1],range[2],xrange[2],range[1],col="grey90",border=NA)
x<-range[1]
repeat{
  lines(c(xrange[1],xrange[2]),c(x,x),col="white")
  x=x+2;if(x>range[2]){break}}

x<- round(xrange[1]+5,-1)
repeat{
  lines(c(x,x),c(range[1],range[2]),col="white")
  x=x+10;if(x>2020){break}}


data=testes_inc_mor;xx<-1
repeat{
smoothingSpline = smooth.spline(data$year, data[,xx], spar=spar, nknots=knot)
lines(smoothingSpline,col=colc[xx],lty=1,lwd=2,lend=1)
xx=xx+1;if(xx>4){break}}

axis(2,las=2,cex.axis=1.4,at=seq(range[1],range[2],by=2),labels=c(range[1],
        rep("",length(seq(range[1],range[2],by=2))-2),range[2]),pos=xrange[1],tck= -0.025)
axis(2,las=2,cex.axis=1.4,at=c(range[1],range[2]),pos=xrange[1],tck= -0.025)
title(ylab="Incidence per 100,000 (ASR - World)", line=0.5, cex.lab=1.4)
axis(side=1,las=1,cex.axis=1.4,at=c(seq(round(xrange[1]+5,-1),2020,by=10)),pos=range[1],
     tck= -0.025)
lines(c(xrange[1],xrange[2]),c(range[1],range[1]))
title(xlab="Year", line=1, cex.lab=1.4,xpd=TRUE)
text(1964,range[2]-0.05*scal,"a",cex=2.5)

xx=1;yy=range[1]+scal*0.25
rect(2000,yy+0.035*scal,2019,yy-0.21*scal,col="white",border="grey50",lwd=0.8)
repeat{
  lines(c(2001,2004),c(yy,yy),lwd=12,col=cola[xx],lend=1)
  xx<-xx+1;yy=yy-(scal*0.058);if(xx>4){break}}


xx=1;yy=range[1]+scal*0.25
text(2011.5,yy,"Denmark",col=cola[xx],cex=1.25);xx=xx+1;yy=yy-(scal*0.058)
text(2011.5,yy,"Finland",col=cola[xx],cex=1.25);xx=xx+1;yy=yy-(scal*0.058)
text(2011.5,yy,"Norway",col=cola[xx],cex=1.25);xx=xx+1;yy=yy-(scal*0.058)
text(2011.4,yy,"Sweden",col=cola[xx],cex=1.25);xx=xx+1;yy=yy-(scal*0.058)


## Plot of prostate cancer ----------------
range<-c(0,120);scal<-range[2]-range[1]
xrange<-c(1961,2020)

plot(NULL,xlim=xrange,ylim=c(range[1],range[2]),xlab="",ylab=""
     ,las=1, axes=FALSE)
rect(xrange[1],range[2],xrange[2],range[1],col="grey90",border=NA)
x<-range[1]
repeat{
  lines(c(xrange[1],xrange[2]),c(x,x),col="white")
  x=x+2;if(x>range[2]){break}}

x<- round(xrange[1]+5,-1)
repeat{
  lines(c(x,x),c(range[1],range[2]),col="white")
  x=x+10;if(x>2020){break}}


data=prostate_inc_mor;xx<-1
repeat{
  smoothingSpline = smooth.spline(data$year, data[,xx], spar=spar, nknots=knot)
  lines(smoothingSpline,col=colc[xx],lty=1,lwd=2,lend=1)
  xx=xx+1;if(xx>4){break}}

axis(2,las=2,cex.axis=1.4,at=seq(range[1],range[2],by=2),labels=c(range[1],
   rep("",length(seq(range[1],range[2],by=2))-2),range[2]),pos=xrange[1],tck= -0.025)
axis(2,las=2,cex.axis=1.4,at=c(range[1],range[2]),pos=xrange[1],tck= -0.025)
#title(ylab="Incidence per 100,000 (ASR - World)", line=0.5, cex.lab=1.4)
axis(side=1,las=1,cex.axis=1.4,at=c(seq(round(xrange[1]+5,-1),2020,by=10)),pos=range[1],
     tck= -0.025)
lines(c(xrange[1],xrange[2]),c(range[1],range[1]))
title(xlab="Year", line=1, cex.lab=1.4,xpd=TRUE)
text(1964,range[2]-0.05*scal,"b",cex=2.5)


## Plot of penis cancer ----------------
range<-c(0,2);scal<-range[2]-range[1]
xrange<-c(1961,2020)

plot(NULL,xlim=xrange,ylim=c(range[1],range[2]),xlab="",ylab=""
     ,las=1, axes=FALSE)
rect(xrange[1],range[2],xrange[2],range[1],col="grey90",border=NA)
x<-range[1]
repeat{
  lines(c(xrange[1],xrange[2]),c(x,x),col="white")
  x=x+2;if(x>range[2]){break}}

x<- round(xrange[1]+5,-1)
repeat{
  lines(c(x,x),c(range[1],range[2]),col="white")
  x=x+10;if(x>2020){break}}

data=penis_inc_mor;xx<-1
repeat{
  smoothingSpline = smooth.spline(data$year, data[,xx], spar=spar, nknots=knot)
  lines(smoothingSpline,col=colc[xx],lty=1,lwd=2,lend=1)
  xx=xx+1;if(xx>4){break}}

axis(2,las=2,cex.axis=1.4,at=seq(range[1],range[2],by=2),labels=c(range[1],
                                                                  rep("",length(seq(range[1],range[2],by=2))-2),range[2]),pos=xrange[1],tck= -0.025)
axis(2,las=2,cex.axis=1.4,at=c(range[1],range[2]),pos=xrange[1],tck= -0.025)
#title(ylab="Incidence per 100,000 (ASR - World)", line=0.5, cex.lab=1.4)
axis(side=1,las=1,cex.axis=1.4,at=c(seq(round(xrange[1]+5,-1),2020,by=10)),pos=range[1],
     tck= -0.025)
lines(c(xrange[1],xrange[2]),c(range[1],range[1]))
title(xlab="Year", line=1, cex.lab=1.4,xpd=TRUE)
text(1964,range[2]-0.05*scal,"c",cex=2.5)


# Plotting - mortality-------------------
## Plot of testes cancer ----------------
range<-c(0,3);scal<-range[2]-range[1]
xrange<-c(1961,2020)

plot(NULL,xlim=xrange,ylim=c(range[1],range[2]),xlab="",ylab=""
     ,las=1, axes=FALSE)
rect(xrange[1],range[2],xrange[2],range[1],col="grey90",border=NA)
x<-range[1]
repeat{
  lines(c(xrange[1],xrange[2]),c(x,x),col="white")
  x=x+2;if(x>range[2]){break}}

x<- round(xrange[1]+5,-1)
repeat{
  lines(c(x,x),c(range[1],range[2]),col="white")
  x=x+10;if(x>2020){break}}


data=testes_inc_mor;xx<-5
repeat{
  smoothingSpline = smooth.spline(data$year, data[,xx], spar=spar, nknots=knot)
  lines(smoothingSpline,col=colc[xx-4],lty=1,lwd=2,lend=1)
  xx=xx+1;if(xx>8){break}}

axis(2,las=2,cex.axis=1.4,at=seq(range[1],range[2],by=2),
     labels=c(range[1],rep("",length(seq(range[1],range[2],by=2))-2),""),pos=xrange[1],tck= -0.025)
axis(2,las=2,cex.axis=1.4,at=c(range[1],range[2]),pos=xrange[1],tck= -0.025)
title(ylab="Mortality per 100,000 (ASR - World)", line=0.5, cex.lab=1.4)
axis(side=1,las=1,cex.axis=1.4,at=c(seq(round(xrange[1]+5,-1),2020,by=10)),pos=range[1],
     tck= -0.025)
lines(c(xrange[1],xrange[2]),c(range[1],range[1]))
title(xlab="Year", line=1, cex.lab=1.4,xpd=TRUE)
text(1964,range[2]-0.05*scal,"d",cex=2.5)

## Plot of prostate cancer ----------------
range<-c(10,24);scal<-range[2]-range[1]
xrange<-c(1961,2020)

plot(NULL,xlim=xrange,ylim=c(range[1],range[2]),xlab="",ylab=""
     ,las=1, axes=FALSE)
rect(xrange[1],range[2],xrange[2],range[1],col="grey90",border=NA)
x<-range[1]
repeat{
  lines(c(xrange[1],xrange[2]),c(x,x),col="white")
  x=x+2;if(x>range[2]){break}}

x<- round(xrange[1]+5,-1)
repeat{
  lines(c(x,x),c(range[1],range[2]),col="white")
  x=x+10;if(x>2020){break}}


data=prostate_inc_mor;xx<-5
repeat{
  smoothingSpline = smooth.spline(data$year, data[,xx], spar=spar, nknots=knot)
  lines(smoothingSpline,col=colc[xx-4],lty=1,lwd=2,lend=1)
  xx=xx+1;if(xx>8){break}}

axis(2,las=2,cex.axis=1.4,at=seq(range[1],range[2],by=2),
     labels=c(range[1],rep("",length(seq(range[1],range[2],by=2))-2),""),pos=xrange[1],tck= -0.025)
axis(2,las=2,cex.axis=1.4,at=c(range[1],range[2]),pos=xrange[1],tck= -0.025)
#title(ylab="Incidence per 100,000 (ASR - World)", line=0.5, cex.lab=1.4)
axis(side=1,las=1,cex.axis=1.4,at=c(seq(round(xrange[1]+5,-1),2020,by=10)),pos=range[1],
     tck= -0.025)
lines(c(xrange[1],xrange[2]),c(range[1],range[1]))
title(xlab="Year", line=1, cex.lab=1.4,xpd=TRUE)
text(1964,range[2]-0.05*scal,"e",cex=2.5)


## Plot of penis cancer ----------------
range<-c(0,0.5);scal<-range[2]-range[1]
xrange<-c(1961,2020)

plot(NULL,xlim=xrange,ylim=c(range[1],range[2]),xlab="",ylab=""
     ,las=1, axes=FALSE)
rect(xrange[1],range[2],xrange[2],range[1],col="grey90",border=NA)
x<-range[1]
repeat{
  lines(c(xrange[1],xrange[2]),c(x,x),col="white")
  x=x+2;if(x>range[2]){break}}

x<- round(xrange[1]+5,-1)
repeat{
  lines(c(x,x),c(range[1],range[2]),col="white")
  x=x+10;if(x>2020){break}}


data=penis_inc_mor;xx<-5
repeat{
  smoothingSpline = smooth.spline(data$year, data[,xx], spar=spar, nknots=knot)
  lines(smoothingSpline,col=colc[xx-4],lty=1,lwd=2,lend=1)
  xx=xx+1;if(xx>8){break}}
axis(2,las=2,cex.axis=1.4,at=c(range[1],range[2]),pos=xrange[1],tck= -0.025)
#title(ylab="Incidence per 100,000 (ASR - World)", line=0.5, cex.lab=1.4)
axis(side=1,las=1,cex.axis=1.4,at=c(seq(round(xrange[1]+5,-1),2020,by=10)),pos=range[1],
     tck= -0.025)
lines(c(xrange[1],xrange[2]),c(range[1],range[1]))
title(xlab="Year", line=1, cex.lab=1.4,xpd=TRUE)
text(1964,range[2]-0.05*scal,"f",cex=2.5)
```




# Non-linear trends in survival


## Defining priors

```{r,errors=FALSE}
prior_group <- c(
  set_prior("normal(0,30)", class = "b", coef = "groupFinland"),
  set_prior("normal(0,30)", class = "b", coef = "groupNorway"),
  set_prior("normal(0,30)", class = "b", coef = "groupSweden"))
```

## Source data

Data were downloaded from [Nordcan](https://nordcan.iarc.fr/en) in October 2022.

### Upload
```{r}
urlfile="https://raw.githubusercontent.com/filip-tichanek/nord_male/main/source-data/testes_1y.csv"
testes_1y<-read.csv(url(urlfile),sep=",")

urlfile="https://raw.githubusercontent.com/filip-tichanek/nord_male/main/source-data/testes_5y.csv"
testes_5y<-read.csv(url(urlfile),sep=",")

urlfile="https://raw.githubusercontent.com/filip-tichanek/nord_male/main/source-data/prostate_1y.csv"
prostate_1y<-read.csv(url(urlfile),sep=",")

urlfile="https://raw.githubusercontent.com/filip-tichanek/nord_male/main/source-data/prostate_5y.csv"
prostate_5y<-read.csv(url(urlfile),sep=",")

urlfile="https://raw.githubusercontent.com/filip-tichanek/nord_male/main/source-data/penis_1y.csv"
penis_1y<-read.csv(url(urlfile),sep=",")

urlfile="https://raw.githubusercontent.com/filip-tichanek/nord_male/main/source-data/penis_5y.csv"
penis_5y<-read.csv(url(urlfile),sep=",")
```


### Data wrangling

```{r}
## testes
x<-2;testes_1y_est<-data.frame(testes_1y[,1])
repeat{testes_1y_est[,x]<-str_sub(testes_1y[,x],1,4);x<-x+1
if(x>5){break}}

repeat{testes_1y_est[,x]<-str_sub(testes_1y[,x-4],6,9);x<-x+1;if(x>9){break}}
x<-2;testes_5y_est<-data.frame(testes_5y[,1])
repeat{testes_5y_est[,x]<-str_sub(testes_5y[,x],1,4);x<-x+1
if(x>5){break}}
repeat{testes_5y_est[,x]<-str_sub(testes_5y[,x-4],6,9);x<-x+1;if(x>9){break}}

testes<-(data.frame(unlist(testes_1y_est[,2:5])));colnames(testes)<-"surv_1y"
testes$cil_1y<-as.numeric(unlist(testes_1y_est[,6:9]))
testes$surv_5y<-as.numeric(unlist(testes_5y_est[,2:5]))
testes$cil_5y<-as.numeric(unlist(testes_5y_est[,6:9]))
testes$year<-rep(seq(1973,2018,by=5),4)
testes$sex<-c(rep("Males",40))
testes$country<-c(rep(c(rep("Denmark",10),rep("Finland",10),rep("Norway",10),rep("Sweden",10)),1))
testes$shou<-c(rep(c(rep("den_",10),rep("fin_",10),rep("nor_",10),rep("swe_",10)),1))
testes$group<-(testes$country)
testes$years10cen<-(testes$year-1995.5)/10
testes$surv_1y<-as.numeric(testes$surv_1y)
testes$surv_5y<-as.numeric(testes$surv_5y)
testes$cil_1y<-as.numeric(testes$cil_1y)
testes$cil_5y<-as.numeric(testes$cil_5y)
testes$se_1y<-(testes$surv_1y-testes$cil_1y)/1.96
testes$se_5y<-(testes$surv_5y-testes$cil_5y)/1.96
testes$surv_cond<-(testes$surv_5y/testes$surv_1y)*100

## prostate
x<-2;prostate_1y_est<-data.frame(prostate_1y[,1])
repeat{prostate_1y_est[,x]<-str_sub(prostate_1y[,x],1,4);x<-x+1
if(x>5){break}}

repeat{prostate_1y_est[,x]<-str_sub(prostate_1y[,x-4],6,9);x<-x+1;if(x>9){break}}
x<-2;prostate_5y_est<-data.frame(prostate_5y[,1])
repeat{prostate_5y_est[,x]<-str_sub(prostate_5y[,x],1,4);x<-x+1
if(x>5){break}}
repeat{prostate_5y_est[,x]<-str_sub(prostate_5y[,x-4],6,9);x<-x+1;if(x>9){break}}

prostate<-(data.frame(unlist(prostate_1y_est[,2:5])));colnames(prostate)<-"surv_1y"
prostate$cil_1y<-as.numeric(unlist(prostate_1y_est[,6:9]))
prostate$surv_5y<-as.numeric(unlist(prostate_5y_est[,2:5]))
prostate$cil_5y<-as.numeric(unlist(prostate_5y_est[,6:9]))
prostate$year<-rep(seq(1973,2018,by=5),4)
prostate$sex<-c(rep("Males",40))
prostate$country<-c(rep(c(rep("Denmark",10),rep("Finland",10),rep("Norway",10),rep("Sweden",10)),1))
prostate$shou<-c(rep(c(rep("den_",10),rep("fin_",10),rep("nor_",10),rep("swe_",10)),1))
prostate$group<-(prostate$country)
prostate$years10cen<-(prostate$year-1995.5)/10
prostate$surv_1y<-as.numeric(prostate$surv_1y)
prostate$surv_5y<-as.numeric(prostate$surv_5y)
prostate$cil_1y<-as.numeric(prostate$cil_1y)
prostate$cil_5y<-as.numeric(prostate$cil_5y)
prostate$se_1y<-(prostate$surv_1y-prostate$cil_1y)/1.96
prostate$se_5y<-(prostate$surv_5y-prostate$cil_5y)/1.96
prostate$surv_cond<-(prostate$surv_5y/prostate$surv_1y)*100

## penis
x<-2;penis_1y_est<-data.frame(penis_1y[,1])
repeat{penis_1y_est[,x]<-str_sub(penis_1y[,x],1,4);x<-x+1
if(x>5){break}}

repeat{penis_1y_est[,x]<-str_sub(penis_1y[,x-4],6,9);x<-x+1;if(x>9){break}}
x<-2;penis_5y_est<-data.frame(penis_5y[,1])
repeat{penis_5y_est[,x]<-str_sub(penis_5y[,x],1,4);x<-x+1
if(x>5){break}}
repeat{penis_5y_est[,x]<-str_sub(penis_5y[,x-4],6,9);x<-x+1;if(x>9){break}}

penis<-(data.frame(unlist(penis_1y_est[,2:5])));colnames(penis)<-"surv_1y"
penis$cil_1y<-as.numeric(unlist(penis_1y_est[,6:9]))
penis$surv_5y<-as.numeric(unlist(penis_5y_est[,2:5]))
penis$cil_5y<-as.numeric(unlist(penis_5y_est[,6:9]))
penis$year<-rep(seq(1973,2018,by=5),4)
penis$sex<-c(rep("Males",40))
penis$country<-c(rep(c(rep("Denmark",10),rep("Finland",10),rep("Norway",10),rep("Sweden",10)),1))
penis$shou<-c(rep(c(rep("den_",10),rep("fin_",10),rep("nor_",10),rep("swe_",10)),1))
penis$group<-(penis$country)
penis$years10cen<-(penis$year-1995.5)/10
penis$surv_1y<-as.numeric(penis$surv_1y)
penis$surv_5y<-as.numeric(penis$surv_5y)
penis$cil_1y<-as.numeric(penis$cil_1y)
penis$cil_5y<-as.numeric(penis$cil_5y)
penis$se_1y<-(penis$surv_1y-penis$cil_1y)/1.96
penis$se_5y<-(penis$surv_5y-penis$cil_5y)/1.96
penis$surv_cond<-(penis$surv_5y/penis$surv_1y)*100

```


## Models

### Fitting models
```{r message=FALSE, warning=FALSE, cache=TRUE, results='hide'}

## testes
set.seed(17)
testes_1y_model<-brm(surv_1y|se(se_1y)~group+s(years10cen,by=group,k=5),
                    family="Gaussian",
                    save_pars = save_pars(all = TRUE),prior=prior_group,seed=17,
                    data=testes,iter=7000, warmup=2000,
                    chains=2,cores=1,control = list(adapt_delta = 0.98))



testes_5y_model<-brm(surv_5y|se(se_5y)~group+s(years10cen,by=group,k=5),
                     family="Gaussian",
                     save_pars= save_pars(all=TRUE),prior=prior_group,seed=17,
                     data=testes,iter=7000,warmup=2000,chains=2,cores=1,
                     control=list(adapt_delta = 0.98))

## prostate
set.seed(17)
prostate_1y_model<-brm(surv_1y|se(se_1y)~group+s(years10cen,by=group,k=5),
                       family="Gaussian",save_pars=save_pars(all=TRUE),
                       prior=prior_group,seed=17,
                       data=prostate,iter=7000,warmup=2000,chains=2,cores=1,
                       control=list(adapt_delta = 0.98))

prostate_5y_model<-brm(surv_5y|se(se_5y)~group+s(years10cen,by=group,k=5),
                       family="Gaussian",save_pars=save_pars(all=TRUE),
                       prior=prior_group,seed=17,data=prostate,
                       iter=7000, warmup=2000,chains=2,cores=1,
                       control = list(adapt_delta = 0.98))

## penis
set.seed(17)
penis_1y_model<-brm(surv_1y|se(se_1y)~group+s(years10cen,by=group,k=5),
                    family="Gaussian",save_pars=save_pars(all=TRUE),
                    prior=prior_group,seed=17,data=penis,
                    iter=7000, warmup=2000,chains=2,cores=1,
                    control = list(adapt_delta = 0.98))

penis_5y_model<-brm(surv_5y|se(se_5y)~group+s(years10cen,by=group,k=5),
                    family="Gaussian",save_pars=save_pars(all=TRUE),
                    prior=prior_group,seed=17,data=penis,
                    iter=7000, warmup=2000,chains=2,cores=1,
                    control = list(adapt_delta = 0.98))
```

## Diagnostics of models

In general, diagnostics uses:
  
[1] *summary(model)*: the columns of intereste are ***Rhat*** (when = 1, model converged well), and columns ***Tail_ESS*** and ***Bulk_ESS*** which should never go under 1,000 and should be mostly above 2,000.  

[2] *pp_check* : graphical tools to explore how well the model predicts real data, as thoroughly described by Gelman [here](http://www.stat.columbia.edu/~gelman/book/). 

You can use also *plot(model)* to see convergence of chains


### testes
```{r message=FALSE, warning=FALSE, results='hide'}
summary(testes_1y_model)
summary(testes_5y_model)

pla<-pp_check(testes_1y_model,type='dens_overlay',ndraws = 100) 
plb<-pp_check(testes_5y_model,type='dens_overlay',ndraws = 100) 
plot_grid(pla, plb, labels=c("A", "B"), ncol = 1, nrow = 2)

pla<-pp_check(testes_1y_model,type='scatter_avg') 
plb<-pp_check(testes_5y_model,type='scatter_avg') 
plot_grid(pla, plb, labels=c("A", "B"), ncol = 1, nrow = 2)

pla<-pp_check(testes_1y_model,type ="stat", stat = "mean")
plb<-pp_check(testes_5y_model,type = "stat", stat = "mean")
plot_grid(pla, plb, labels=c("A", "B"), ncol = 1, nrow = 2)

pla<-pp_check(testes_1y_model,type="stat_2d", stat = c("max", "min"),ndraws=200)
plb<-pp_check(testes_5y_model,type="stat_2d", stat = c("max", "min"),ndraws=200)
plot_grid(pla, plb, labels=c("A", "B"), ncol = 1, nrow = 2)
```

### prostate
```{r message=FALSE, warning=FALSE, results='hide'}
summary(prostate_1y_model)
summary(prostate_5y_model)
pla<-pp_check(prostate_1y_model,type='dens_overlay',ndraws = 100) 
plb<-pp_check(prostate_5y_model,type='dens_overlay',ndraws = 100) 
plot_grid(pla, plb, labels=c("A", "B"), ncol = 1, nrow = 2)

pla<-pp_check(prostate_1y_model,type='scatter_avg') 
plb<-pp_check(prostate_5y_model,type='scatter_avg') 
plot_grid(pla, plb, labels=c("A", "B"), ncol = 1, nrow = 2)

pla<-pp_check(prostate_1y_model,type ="stat", stat = "mean")
plb<-pp_check(prostate_5y_model,type = "stat", stat = "mean")
plot_grid(pla, plb, labels=c("A", "B"), ncol = 1, nrow = 2)

pla<-pp_check(prostate_1y_model,type="stat_2d", stat = c("max", "min"),ndraws=200)
plb<-pp_check(prostate_5y_model,type="stat_2d", stat = c("max", "min"),ndraws=200)
plot_grid(pla, plb, labels=c("A", "B"), ncol = 1, nrow = 2)
```

### penis
```{r message=FALSE, warning=FALSE, results='hide'}
summary(prostate_1y_model)
summary(prostate_5y_model)
pla<-pp_check(penis_1y_model,type='dens_overlay',ndraws = 100) 
plb<-pp_check(penis_5y_model,type='dens_overlay',ndraws = 100) 
plot_grid(pla, plb, labels=c("A", "B"), ncol = 1, nrow = 2)

pla<-pp_check(penis_1y_model,type='scatter_avg') 
plb<-pp_check(penis_5y_model,type='scatter_avg') 
plot_grid(pla, plb, labels=c("A", "B"), ncol = 1, nrow = 2)

pla<-pp_check(penis_1y_model,type ="stat", stat = "mean")
plb<-pp_check(penis_5y_model,type = "stat", stat = "mean")
plot_grid(pla, plb, labels=c("A", "B"), ncol = 1, nrow = 2)

pla<-pp_check(penis_1y_model,type="stat_2d", stat = c("max", "min"),ndraws=200)
plb<-pp_check(penis_5y_model,type="stat_2d", stat = c("max", "min"),ndraws=200)
plot_grid(pla, plb, labels=c("A", "B"), ncol = 1, nrow = 2)
```

## Functions for plotting

All the functions below serve for plotting of nonlinear trends in survival and related uncertainty. All the functions below use extracted posterior samples as an input *data*. The samples represent estimation of survival over the 50 years (1971 to 2020) for specific country, cancer, and type of survival (1-year, 5-years, conditional [5y/1y]. The time period of 50 years must be divided to sequence of 500 numbers (1/10 of a year). 

### ***breakpo*** - detection of breaking points

Function serves to identify *breaking points*, i.e. times when the annual change of survival changed with at least 95% plausibility. This was assessed by calculation of the 2nd derivation of the given survival measure and its 95% CI; the ‘breaking point’ was defined as a peak value within at least a 3-year interval where 95% CI for the 2nd derivation did not cross zero. If the 2nd derivation is plausibly non-zero for at least 3 years, the function takes the peak in the 2nd derivation (within the identified time interval) as the *breaking point* (must be between the years 1976 and 2016)

There is one additional argument 'arb' which should be zero except for the situation when the multiple breaking points overlap. The argument only move the breaking points by given value to avoid overlapping. 

The function often gives warning: *## Error in xy.coords(x, y): 'x' and 'y' lengths differ*. Warning occurs when there is no *breaking point* in a given group and survival measure, ie, it is not real error. 

```{r message=FALSE, warning=FALSE}
breakpo<-function(data,arb){ 
  data<-data.frame((data[,-1] - data[,-ncol(data)])*10)
  data<-data.frame(data[,-1] - data[,-ncol(data)])
  data=sapply(data, function(p) quantile(p, probs = c(0.025,0.975,0.5)))
  cbinl<-c()
  x=1
  repeat{
    cbinl[x]<-
      if(data[1,x]>0|data[2,x]<0){cbinl[x]=1}else{cbinl[x]=0}
    x<-x+1
    if(x>length(data[1,])){break}}
  cbin=1;x<-1
  repeat{
    cbin[x+1]<-abs(cbinl[x]-cbinl[x+1])
    x=x+1
    if(x>(length(cbinl)-1)){break}}
  data<-data.frame(rbind(data,cbin));data[5,]<-yreal[c(2:499)]-0.049
  data[6,]<-1:498;data[4,51]<-1;data[4,449]<-1
  
  row.names(data)<-c("cil","ciu","est","stat_change","year","timepoint")
  data2=t(data[,data[4,]==1])
  data2<-data.frame(data2)
  tr<-subset(data2,data2$timepoint>50&data2$timepoint<450)
  y=1
  bp<-c()
  bx<-1
  repeat{
    if( (tr[y,1]<0) & (tr[y,2]<0) & (  (tr$year[y+1]-tr$year[y])>3   )  ) {
      tr2<-data[,tr[y,6]:(tr[y+1,6]-2)]
      bp[bx]<- tr2[,order(tr2[3,],decreasing=F)[1]][5]
      bx=bx+1
    }
    if( (tr[y,1]>0) & (tr[y,2]>0) & (  (tr$year[y+1]-tr$year[y])>3   )) {
      tr2<-data[,tr[y,6]:(tr[y+1,6]-2)]
      bp[bx]<- tr2[,order(tr2[3,],decreasing=T)[1]][5]
      bx=bx+1
    }
    y=y+1;if(y>(dim(tr)[1]-1)){break}}
  y<-1;repeat{
    lines(c(bp[y]+arb,bp[y]+arb),c(range[1],range[1]+0.025*scal),col=cola[xx],lwd=3.5,lend=1)
    lines(c(bp[y]+arb,bp[y]+arb),c(range[2],range[2]-0.025*scal),col=cola[xx],lwd=3.5,lend=1)
    lines(c(bp[y]+arb,bp[y]+arb),c(range[1],range[1]+0.999*scal),col=colc[xx],lwd=1,lend=1,lty=2)
    y=y+1;if(y>length(bp)){break}}
  print(bp)
  print(tr)}
```


### ***polyg_surv*** 
for drawing 95% credible interval for survival indicators

```{r message=FALSE, warning=FALSE}
polyg_surv<-function(data){ 
  data<-data.frame(data)
  data<-sapply(data, function(p) quantile(p, probs = c(0.025,0.975,0.5)))
  cis<-c(data[1,],rev(data[2,]))
  x<-c(yreal[1:500],yreal[500:1])
  cis[cis<range[1]]<-range[1]
  cis[cis>range[2]]<-range[2]
  polygon(x,cis,border=NA,col=colb[xx],xpd=F)}
```

### ***surv_fit*** 
Fit curve of survival trend over the 50 years. Solid lines imply that that the curve is increasing or decreasing with at least 95% plausibility (95% credible interval for 1st derivation of estimated survival does not cross zero) for at least 5 years. Dashed lines show otherwise (no detectable change in the survival)

```{r message=FALSE, warning=FALSE}
surv_fit<-function(dat2){
  dat2=data.frame(dat2)
  cis= data.frame((dat2[,-1] - dat2[,-ncol(dat2)])*10)
  cis=sapply(cis, function(p) quantile(p, probs = c(0.025,0.975)))
  est=sapply(dat2, function(p) quantile(p, probs = c(0.5)))
  data=cis
  cbinl<-c()
  x=1
  repeat{
    cbinl[x]<-
      if(data[1,x]>0|data[2,x]<0){cbinl[x]=1}else{cbinl[x]=0}
    x<-x+1
    if(x>length(data[1,])){break}}
  cbin=1;x<-1
  repeat{
    cbin[x+1]<-abs(cbinl[x]-cbinl[x+1])
    x=x+1
    if(x>(length(cbinl)-1)){break}}
  data<-data.frame(rbind(data,cbin));data[4,]<-yreal[c(2:500)]-0.049;data[5,]<-1:499
  row.names(data)<-c("cil","ciu","stat_change","year","timepoint");data[3,499]<-1
  data2=t(data[,data[3,]==1])
  data2<-data.frame(data2)
  tr<-data2;y=1
  yreal=seq(1971,2020,length=500)
  repeat{
    if((((tr[y,1]<0)&(tr[y,2])<0))|(((tr[y,1]>0)&(tr[y,2])>0))&(tr[y+1,4]-tr[y,4])>5)
    {lines(yreal[tr[y,5]:tr[y+1,5]],
           est[tr[y,5]:tr[y+1,5]],lwd=1.9,col=cola[xx],lend=1)}
    y<-y+1;if(y>dim(tr)[1]-1){break}}
  lines(yreal[1:500],est,lwd=1,col=cola[xx],lty=2,lend=1)}
```

### ***polyg_slope*** 
Draws 95% credible interval for slope of the survival trend (1st derivation of the estimated survival trend)

```{r message=FALSE, warning=FALSE}
polyg_slope<-function(data){ 
  x<-c(yreal[1:499],yreal[499:1])
  data<-data.frame((data[,-1] - data[,-ncol(data)]))*10
  data=sapply(data, function(p) quantile(p, probs = c(0.025,0.975,0.5)))
  cis<-c(data[1,],rev(data[2,]))
  cis[cis<range[1]]<-range[1]
  cis[cis>range[2]]<-range[2]
  polygon(x,cis,border=NA,col=colb[xx])}
```

### ***slope_fit*** 
Fit curve of slope of the survival trend over the 50 years. Solid lines imply that that the curve is increasing or decreasing with at least 95% plausibility (95% credible interval for 2st derivation of estimated survival does not cross zero) for at least 3 years. Dashed lines show otherwise (no detectable change in the slope of the survival trend)

```{r message=FALSE, warning=FALSE}
slope_fit<-function(data){
  data = data.frame((data[,-1] - data[,-ncol(data)])*10)
  dar2<-sapply(data, function(p) quantile(p, probs = c(0.5)))
  data<-data.frame(data[,-1] - data[,-ncol(data)])
  data=sapply(data, function(p) quantile(p, probs = c(0.025,0.975,0.5)))
  cbinl<-c()
  x=1
  repeat{
    cbinl[x]<-
      if(data[1,x]>0|data[2,x]<0){cbinl[x]=1}else{cbinl[x]=0}
    x<-x+1
    if(x>length(data[1,])){break}}
  cbin=1;x<-1
  repeat{
    cbin[x+1]<-abs(cbinl[x]-cbinl[x+1])
    x=x+1
    if(x>(length(cbinl)-1)){break}}
  data<-data.frame(rbind(data,cbin));data[5,]<-yreal[c(2:499)]-0.049;data[6,]<-1:498
  row.names(data)<-c("cil","ciu","est","stat_change","year","timepoint")
  data[4,450]<-1
  data[4,50]<-1
  data=data[,50:450]
  lines(yreal[1:499],dar2,lwd=1,col=cola[xx],lty=2,lend=1)
  data2=(t(data[,data[4,]==1]))
  data2<-data.frame(data2)
  tr<-data2;y=1
  yreal=seq(1971,2020,length=498)
  repeat{
    if(((((tr[y,1]<0)&(tr[y,2])<0))|(((tr[y,1]>0)&(tr[y,2])>0)))&(tr[y+1,5]-tr[y,5])>3)
    {lines(yreal[tr[y,6]:tr[y+1,6]],
           dar2[tr[y,6]:tr[y+1,6]]
           ,lwd=1.9,col=cola[xx],lend=1)}
    y<-y+1;if(y>dim(tr)[1]-1){break}}
  return(data2)}
```
  
## Extraction of posterior samples

```{r message=FALSE, warning=FALSE}
yreal <- seq(1971,2020,length=500)
first <- expand.grid(years10cen = ((yreal-1995.5)/10),
                     group = c('Denmark','Finland','Norway','Sweden'), y = 0)

ms_1y<-posterior_smooths(testes_1y_model,smooth="s(years10cen,by=group,k=5)",newdata = first)
post_fix<-as.data.frame(testes_1y_model, variable = c("b_Intercept","b_groupFinland",
      "b_groupNorway","b_groupSweden"))

post_testes_den_1y<-ms_1y[,1:500]     +post_fix[,1]
post_testes_fin_1y<-ms_1y[,501:1000]  +post_fix[,1]+post_fix[,2]
post_testes_nor_1y<-ms_1y[,1001:1500] +post_fix[,1]+post_fix[,3]
post_testes_swe_1y<-ms_1y[,1501:2000] +post_fix[,1]+post_fix[,4]

ms_5y<-posterior_smooths(testes_5y_model,smooth="s(years10cen,by=group,k=5)",newdata = first)
post_fix<-as.data.frame(testes_5y_model, variable = c("b_Intercept","b_groupFinland",
                                                      "b_groupNorway","b_groupSweden"))
post_testes_den_5y<-ms_5y[,1:500]     +post_fix[,1]
post_testes_fin_5y<-ms_5y[,501:1000]  +post_fix[,1]+post_fix[,2]
post_testes_nor_5y<-ms_5y[,1001:1500] +post_fix[,1]+post_fix[,3]
post_testes_swe_5y<-ms_5y[,1501:2000] +post_fix[,1]+post_fix[,4]

post_testes_den_cond<-(post_testes_den_5y/post_testes_den_1y)*100
post_testes_fin_cond<-(post_testes_fin_5y/post_testes_fin_1y)*100
post_testes_nor_cond<-(post_testes_nor_5y/post_testes_nor_1y)*100
post_testes_swe_cond<-(post_testes_swe_5y/post_testes_swe_1y)*100


## prostate posterior extraction ---------------------------------------------------
ms_1y<-posterior_smooths(prostate_1y_model,smooth="s(years10cen,by=group,k=5)",newdata = first)
post_fix<-as.data.frame(prostate_1y_model, variable = c("b_Intercept","b_groupFinland",
                                                      "b_groupNorway","b_groupSweden"))
post_prostate_den_1y<-ms_1y[,1:500]     +post_fix[,1]
post_prostate_fin_1y<-ms_1y[,501:1000]  +post_fix[,1]+post_fix[,2]
post_prostate_nor_1y<-ms_1y[,1001:1500] +post_fix[,1]+post_fix[,3]
post_prostate_swe_1y<-ms_1y[,1501:2000] +post_fix[,1]+post_fix[,4]

ms_5y<-posterior_smooths(prostate_5y_model,smooth="s(years10cen,by=group,k=5)",newdata = first)
post_fix<-as.data.frame(prostate_5y_model, variable = c("b_Intercept","b_groupFinland",
                                                      "b_groupNorway","b_groupSweden"))
post_prostate_den_5y<-ms_5y[,1:500]     +post_fix[,1]
post_prostate_fin_5y<-ms_5y[,501:1000]  +post_fix[,1]+post_fix[,2]
post_prostate_nor_5y<-ms_5y[,1001:1500] +post_fix[,1]+post_fix[,3]
post_prostate_swe_5y<-ms_5y[,1501:2000] +post_fix[,1]+post_fix[,4]

post_prostate_den_cond<-(post_prostate_den_5y/post_prostate_den_1y)*100
post_prostate_fin_cond<-(post_prostate_fin_5y/post_prostate_fin_1y)*100
post_prostate_nor_cond<-(post_prostate_nor_5y/post_prostate_nor_1y)*100
post_prostate_swe_cond<-(post_prostate_swe_5y/post_prostate_swe_1y)*100


## penis posterior extraction ---------------------------------------------------
ms_1y<-posterior_smooths(penis_1y_model,smooth="s(years10cen,by=group,k=5)",newdata = first)
post_fix<-as.data.frame(penis_1y_model, variable = c("b_Intercept","b_groupFinland",
                                                      "b_groupNorway","b_groupSweden"))
post_penis_den_1y<-ms_1y[,1:500]     +post_fix[,1]
post_penis_fin_1y<-ms_1y[,501:1000]  +post_fix[,1]+post_fix[,2]
post_penis_nor_1y<-ms_1y[,1001:1500] +post_fix[,1]+post_fix[,3]
post_penis_swe_1y<-ms_1y[,1501:2000] +post_fix[,1]+post_fix[,4]

ms_5y<-posterior_smooths(penis_5y_model,smooth="s(years10cen,by=group,k=5)",newdata = first)
post_fix<-as.data.frame(penis_5y_model, variable = c("b_Intercept","b_groupFinland",
                                                      "b_groupNorway","b_groupSweden"))
post_penis_den_5y<-ms_5y[,1:500]     +post_fix[,1]
post_penis_fin_5y<-ms_5y[,501:1000]  +post_fix[,1]+post_fix[,2]
post_penis_nor_5y<-ms_5y[,1001:1500] +post_fix[,1]+post_fix[,3]
post_penis_swe_5y<-ms_5y[,1501:2000] +post_fix[,1]+post_fix[,4]

post_penis_den_cond<-(post_penis_den_5y/post_penis_den_1y)*100
post_penis_fin_cond<-(post_penis_fin_5y/post_penis_fin_1y)*100
post_penis_nor_cond<-(post_penis_nor_5y/post_penis_nor_1y)*100
post_penis_swe_cond<-(post_penis_swe_5y/post_penis_swe_1y)*100
```

## Figures

### Denmark

```{r, fig.height=5.75,fig.width=8.13, fig.cap="Fig. 1. Relative 1- , 5/1- (conditional) and 5-years survival in Danish men with (a) testicular, (b) prostate and (c) penile cancer. The vertical lines mark a significant change in the survival trends (‘breaking points’) and the bottom curves show estimated annual changes in survival. The curves are solid if there is >95% plausibility of the growth or decline. Shadow areas indicate 95% credible interval. All curves are color coded (see the insert)",error=TRUE,warning=FALSE,results=FALSE,message=FALSE}
## DEN figures -----------------
m <- matrix(c(1,2,3
             ,4,6,8
             ,5,7,9), nrow = 3, ncol =3 ,byrow = TRUE)
layout(mat = m,heights = c(0.03,0.6,0.37))
par(mgp=c(1.6,0.62,0))
par(mar=c(0,0,0,0))


plot(NULL, axes=FALSE,xlab="",ylab="",xlim=c(-1,1),ylim=c(-0.85,0.85))
text(0,-0.2,"Testicular cancer",cex=1.6,font=3,xpd=TRUE)

plot(NULL, axes=FALSE,xlab="",ylab="",xlim=c(-1,1),ylim=c(-0.85,0.85))
text(0,-0.2,"Prostate cancer",cex=1.6,font=3,xpd=TRUE)

plot(NULL, axes=FALSE,xlab="",ylab="",xlim=c(-1,1),ylim=c(-0.85,0.85))
text(0,-0.2,"Penile cancer",cex=1.6,font=3,xpd=TRUE)

range_b<-c(30,110);scal_b<-range_b[2]-range_b[1]
range_c<-c(-1,4);scal_c<-range_c[2]-range_c[1]

#
### testes  % ------------------------------------------------------
range<-range_b;scal=scal_b
par(mar=c(0,2.1,0,0))

plot(NULL,xlim=c(1971,2020),ylim=c(range[1],range[2]),xlab="",ylab=""
     ,las=1, axes=FALSE)
rect(1971,range[2],2020,range[1],col="grey90",border=NA)
x<-10
repeat{
  lines(c(1971,2020),c(x,x),col="white")
  x=x+10;if(x>100){break}}

x<-1980
repeat{
  lines(c(x,x),c(range[1],range[2]),col="white")
  x=x+10;if(x>2020){break}}

# breaking points identification
xx=1
breakpo(post_testes_den_1y,0)
xx=xx+1
breakpo(post_testes_den_cond,0)
xx=xx+1
breakpo(post_testes_den_5y,0)

# 95% credible interval for survival
xx=1
polyg_surv(post_testes_den_1y);xx=xx+1
polyg_surv(post_testes_den_cond);xx=xx+1
polyg_surv(post_testes_den_5y)

# data points of estimated survival
xx=1
points(testes[testes$shou=="den_",]$surv_1y~testes[testes$shou=="den_",]$year
       ,pch=17,col=colc[xx],cex=1);xx=xx+1
points(testes[testes$shou=="den_",]$surv_cond~testes[testes$shou=="den_",]$year
       ,pch=17,col=colc[xx],cex=1);xx=xx+1
points(testes[testes$shou=="den_",]$surv_5y~testes[testes$shou=="den_",]$year
       ,pch=17,col=colc[xx],cex=1)

# fitted lines for survival
xx=1
surv_fit(post_testes_den_1y);xx=xx+1
surv_fit(post_testes_den_cond);xx=xx+1
surv_fit(post_testes_den_5y);xx=xx+1

axis(2,las=2,cex.axis=1.4,at=seq(range[1],range[2],by=10),
     labels=c(range[1],rep("",length(seq(range[1],range[2],by=10))-3),100,""),
     pos=1971)
title(ylab="Relative survival (%)", line=0.5, cex.lab=1.5,adj= 0.4)
axis(side=1,las=1,cex.axis=1.4,at=c(seq(1980,2020,by=10)),pos=range[1],
     labels = c(rep("",5)))
lines(c(1971,2020),c(range[1],range[1]))
      
# legend
xpo=0
ypo=0.2
xx=1;yy=range[1]+scal*ypo
rect(1996+xpo,yy+0.035*scal,2017.9+xpo,yy-0.15*scal,col="white",border="grey50",lwd=0.8)
repeat{
  points(2000+xpo,yy,pch=17,col=cola[xx],cex=1.2)
  lines(c(1997+xpo,2003+xpo),c(yy,yy),lwd=1.6,col=cola[xx],lend=1)
  xx<-xx+1;yy=yy-(scal*0.058);if(xx>3){break}}
xx=1;yy=range[1]+scal*ypo
text(2010+xpo,yy,"1-year",col=cola[xx],cex=1.25);xx=xx+1;yy=yy-(scal*0.058)
text(2010+xpo,yy,"5/1-year",col=cola[xx],cex=1.25);xx=xx+1;yy=yy-(scal*0.058)
text(2010+xpo,yy,"5-year",col=cola[xx],cex=1.25);xx=xx+1;yy=yy-(scal*0.058)
text(1974,range[2]-0.05*scal,"a",cex=2.2)

#### slope plot --------------
range<-range_c;scal=scal_c
par(mar=c(2.5,2.1,0,0))

plot(NULL,xlim=c(1971,2020),ylim=c(range[1],range[2]),xlab="",ylab=""
     ,las=1, axes=FALSE)
rect(1971,range[2],2020,range[1],col="grey90",border=NA)
x<-range[1]
repeat{
  lines(c(1971,2020),c(x,x),col="white")
  x=x+1;if(x>100){break}}

x<-1980
repeat{
  lines(c(x,x),c(range[1],range[2]),col="white")
  x=x+10;if(x>2020){break}}

xx=1
polyg_slope(post_testes_den_1y);xx=xx+1
polyg_slope(post_testes_den_cond);xx=xx+1
polyg_slope(post_testes_den_5y)

xx=1
slope_fit(post_testes_den_1y);xx=xx+1
slope_fit(post_testes_den_cond);xx=xx+1
slope_fit(post_testes_den_5y)

axis(2,las=2,cex.axis=1.4,at=c(seq(range[1],range[2],by=1)),
     labels=c(range[1],rep("",length(seq(range[1],range[2],by=1))-2),range[2]),pos=1971)
lines(c(1971,2020),c(0,0),col="grey50")

title(ylab=expression(paste(delta, " (survival)")), line=0.5, cex.lab=1.5,adj= 0.4)

axis(side=1,las=1,cex.axis=1.4,at=c(seq(1980,2020,by=10)),pos=range[1],labels=c(rep("",5)))
text(c(seq(1978,2020,by=20)),c(rep(range[1]-0.085*scal,3)),c(seq(1980,2020,by=20)),xpd=TRUE,cex=1.4);
lines(c(1971,2020),c(range[1],range[1]))
title(xlab="Year", line=1.4, cex.lab=1.4,xpd=TRUE)


### prostate  % ------------------------------------------------------
range<-range_b;scal=scal_b
par(mar=c(0,2.1,0,0))

plot(NULL,xlim=c(1971,2020),ylim=c(range[1],range[2]),xlab="",ylab=""
     ,las=1, axes=FALSE)
rect(1971,range[2],2020,range[1],col="grey90",border=NA)
x<-10
repeat{
  lines(c(1971,2020),c(x,x),col="white")
  x=x+10;if(x>100){break}}

x<-1980
repeat{
  lines(c(x,x),c(range[1],range[2]),col="white")
  x=x+10;if(x>2020){break}}

# breaking points identification
xx=1
breakpo(post_prostate_den_1y,-1)
xx=xx+1
breakpo(post_prostate_den_cond,0)
xx=xx+1
breakpo(post_prostate_den_5y,1)

# 95% credible interval for survival
xx=1
polyg_surv(post_prostate_den_1y);xx=xx+1
polyg_surv(post_prostate_den_cond);xx=xx+1
polyg_surv(post_prostate_den_5y)

# data points of estimated survival
xx=1
points(prostate[prostate$shou=="den_",]$surv_1y~prostate[prostate$shou=="den_",]$year
       ,pch=17,col=colc[xx],cex=1);xx=xx+1
points(prostate[prostate$shou=="den_",]$surv_cond~prostate[prostate$shou=="den_",]$year
       ,pch=17,col=colc[xx],cex=1);xx=xx+1
points(prostate[prostate$shou=="den_",]$surv_5y~prostate[prostate$shou=="den_",]$year
       ,pch=17,col=colc[xx],cex=1)

# fitted lines for survival
xx=1
surv_fit(post_prostate_den_1y);xx=xx+1
surv_fit(post_prostate_den_cond);xx=xx+1
surv_fit(post_prostate_den_5y);xx=xx+1

axis(2,las=2,cex.axis=1.4,at=seq(range[1],range[2],by=10),
     labels=c(range[1],rep("",length(seq(range[1],range[2],by=10))-3),100,""),
     pos=1971)
axis(side=1,las=1,cex.axis=1.4,at=c(seq(1980,2020,by=10)),pos=range[1],
     labels = c(rep("",5)))
lines(c(1971,2020),c(range[1],range[1]))
text(1974,range[2]-0.05*scal,"b",cex=2.2)

#### slope plot --------------
range<-range_c;scal=scal_c
par(mar=c(2.5,2.1,0,0))

plot(NULL,xlim=c(1971,2020),ylim=c(range[1],range[2]),xlab="",ylab=""
     ,las=1, axes=FALSE)
rect(1971,range[2],2020,range[1],col="grey90",border=NA)
x<-range[1]
repeat{
  lines(c(1971,2020),c(x,x),col="white")
  x=x+1;if(x>100){break}}

x<-1980
repeat{
  lines(c(x,x),c(range[1],range[2]),col="white")
  x=x+10;if(x>2020){break}}

xx=1
polyg_slope(post_prostate_den_1y);xx=xx+1
polyg_slope(post_prostate_den_cond);xx=xx+1
polyg_slope(post_prostate_den_5y)

xx=1
slope_fit(post_prostate_den_1y);xx=xx+1
slope_fit(post_prostate_den_cond);xx=xx+1
slope_fit(post_prostate_den_5y)

axis(2,las=2,cex.axis=1.4,at=c(seq(range[1],range[2],by=1)),
     labels=c(range[1],rep("",length(seq(range[1],range[2],by=1))-2),range[2]),pos=1971)
lines(c(1971,2020),c(0,0),col="grey50")

axis(side=1,las=1,cex.axis=1.4,at=c(seq(1980,2020,by=10)),pos=range[1],labels=c(rep("",5)))
text(c(seq(1978,2020,by=20)),c(rep(range[1]-0.085*scal,3)),c(seq(1980,2020,by=20)),xpd=TRUE,cex=1.4);
lines(c(1971,2020),c(range[1],range[1]))
title(xlab="Year", line=1.4, cex.lab=1.4,xpd=TRUE)


### penis  % ------------------------------------------------------
range<-range_b;scal=scal_b
par(mar=c(0,2.1,0,0))

plot(NULL,xlim=c(1971,2020),ylim=c(range[1],range[2]),xlab="",ylab=""
     ,las=1, axes=FALSE)
rect(1971,range[2],2020,range[1],col="grey90",border=NA)
x<-10
repeat{
  lines(c(1971,2020),c(x,x),col="white")
  x=x+10;if(x>100){break}}

x<-1980
repeat{
  lines(c(x,x),c(range[1],range[2]),col="white")
  x=x+10;if(x>2020){break}}

# breaking points identification
xx=1
breakpo(post_penis_den_1y,-1)
xx=xx+1
breakpo(post_penis_den_cond,0)
xx=xx+1
breakpo(post_penis_den_5y,1)

# 95% credible interval for survival
xx=1
polyg_surv(post_penis_den_1y);xx=xx+1
polyg_surv(post_penis_den_cond);xx=xx+1
polyg_surv(post_penis_den_5y)

# data points of estimated survival
xx=1
points(penis[penis$shou=="den_",]$surv_1y~penis[penis$shou=="den_",]$year
       ,pch=17,col=colc[xx],cex=1);xx=xx+1
points(penis[penis$shou=="den_",]$surv_cond~penis[penis$shou=="den_",]$year
       ,pch=17,col=colc[xx],cex=1);xx=xx+1
points(penis[penis$shou=="den_",]$surv_5y~penis[penis$shou=="den_",]$year
       ,pch=17,col=colc[xx],cex=1)

# fitted lines for survival
xx=1
surv_fit(post_penis_den_1y);xx=xx+1
surv_fit(post_penis_den_cond);xx=xx+1
surv_fit(post_penis_den_5y);xx=xx+1

axis(2,las=2,cex.axis=1.4,at=seq(range[1],range[2],by=10),
     labels=c(range[1],rep("",length(seq(range[1],range[2],by=10))-3),100,""),
     pos=1971)
axis(side=1,las=1,cex.axis=1.4,at=c(seq(1980,2020,by=10)),pos=range[1],
     labels = c(rep("",5)))
lines(c(1971,2020),c(range[1],range[1]))
text(1974,range[2]-0.05*scal,"c",cex=2.2)

ypo=-60
xfl=48
rect(1956+xfl,range[1]+scal*0.965+ypo,1968+xfl,range[1]+scal*0.84+ypo,col="red2",border="grey50")
rect(1956+xfl,range[1]+scal*0.9135+ypo,1968+xfl,range[1]+scal*0.8865+ypo,col="white",border=NA)
rect(1960+xfl,range[1]+scal*0.965+ypo,1962+xfl,range[1]+scal*0.84+ypo,col="white",border=NA)

#### slope plot --------------
range<-range_c;scal=scal_c
par(mar=c(2.5,2.1,0,0))

plot(NULL,xlim=c(1971,2020),ylim=c(range[1],range[2]),xlab="",ylab=""
     ,las=1, axes=FALSE)
rect(1971,range[2],2020,range[1],col="grey90",border=NA)
x<-range[1]
repeat{
  lines(c(1971,2020),c(x,x),col="white")
  x=x+1;if(x>100){break}}

x<-1980
repeat{
  lines(c(x,x),c(range[1],range[2]),col="white")
  x=x+10;if(x>2020){break}}

xx=1
polyg_slope(post_penis_den_1y);xx=xx+1
polyg_slope(post_penis_den_cond);xx=xx+1
polyg_slope(post_penis_den_5y)

xx=1
slope_fit(post_penis_den_1y);xx=xx+1
slope_fit(post_penis_den_cond);xx=xx+1
slope_fit(post_penis_den_5y)

axis(2,las=2,cex.axis=1.4,at=c(seq(range[1],range[2],by=1)),
     labels=c(range[1],rep("",length(seq(range[1],range[2],by=1))-2),range[2]),pos=1971)
lines(c(1971,2020),c(0,0),col="grey50")

axis(side=1,las=1,cex.axis=1.4,at=c(seq(1980,2020,by=10)),pos=range[1],labels=c(rep("",5)))
text(c(seq(1978,2020,by=20)),c(rep(range[1]-0.085*scal,3)),c(seq(1980,2020,by=20)),xpd=TRUE,cex=1.4);
lines(c(1971,2020),c(range[1],range[1]))
title(xlab="Year", line=1.4, cex.lab=1.4,xpd=TRUE)
```

### Finland

```{r, fig.height=5.75,fig.width=8.13, fig.cap="Fig. 2. Relative 1- , 5/1- (conditional) and 5-years survival in Finnish men with (a) testicular, (b) prostate and (c) penile cancer. The vertical lines mark a significant change in the survival trends (‘breaking points’) and the bottom curves show estimated annual changes in survival. The curves are solid if there is >95% plausibility of the growth or decline. Shadow areas indicate 95% credible interval. All curves are color coded (see the insert)",error=TRUE,warning=FALSE,results=FALSE,message=FALSE}
## FIN figures -----------------
m <- matrix(c(1,2,3
              ,4,6,8
              ,5,7,9), nrow = 3, ncol =3 ,byrow = TRUE)
layout(mat = m,heights = c(0.03,0.6,0.37))
par(mgp=c(1.6,0.62,0))
par(mar=c(0,0,0,0))


plot(NULL, axes=FALSE,xlab="",ylab="",xlim=c(-1,1),ylim=c(-0.85,0.85))
text(0,-0.2,"Testicular cancer",cex=1.6,font=3,xpd=TRUE)

plot(NULL, axes=FALSE,xlab="",ylab="",xlim=c(-1,1),ylim=c(-0.85,0.85))
text(0,-0.2,"Prostate cancer",cex=1.6,font=3,xpd=TRUE)

plot(NULL, axes=FALSE,xlab="",ylab="",xlim=c(-1,1),ylim=c(-0.85,0.85))
text(0,-0.2,"Penile cancer",cex=1.6,font=3,xpd=TRUE)

range_b<-c(30,110);scal_b<-range_b[2]-range_b[1]
range_c<-c(-1,4);scal_c<-range_c[2]-range_c[1]

#
### testes  % ------------------------------------------------------
range<-range_b;scal=scal_b
par(mar=c(0,2.1,0,0))

plot(NULL,xlim=c(1971,2020),ylim=c(range[1],range[2]),xlab="",ylab=""
     ,las=1, axes=FALSE)
rect(1971,range[2],2020,range[1],col="grey90",border=NA)
x<-10
repeat{
  lines(c(1971,2020),c(x,x),col="white")
  x=x+10;if(x>100){break}}

x<-1980
repeat{
  lines(c(x,x),c(range[1],range[2]),col="white")
  x=x+10;if(x>2020){break}}

# breaking points identification
xx=1
breakpo(post_testes_fin_1y,0)
xx=xx+1
breakpo(post_testes_fin_cond,0)
xx=xx+1
breakpo(post_testes_fin_5y,0)

# 95% credible interval for survival
xx=1
polyg_surv(post_testes_fin_1y);xx=xx+1
polyg_surv(post_testes_fin_cond);xx=xx+1
polyg_surv(post_testes_fin_5y)

# data points of estimated survival
xx=1
points(testes[testes$shou=="fin_",]$surv_1y~testes[testes$shou=="fin_",]$year
       ,pch=17,col=colc[xx],cex=1);xx=xx+1
points(testes[testes$shou=="fin_",]$surv_cond~testes[testes$shou=="fin_",]$year
       ,pch=17,col=colc[xx],cex=1);xx=xx+1
points(testes[testes$shou=="fin_",]$surv_5y~testes[testes$shou=="fin_",]$year
       ,pch=17,col=colc[xx],cex=1)

# fitted lines for survival
xx=1
surv_fit(post_testes_fin_1y);xx=xx+1
surv_fit(post_testes_fin_cond);xx=xx+1
surv_fit(post_testes_fin_5y);xx=xx+1

axis(2,las=2,cex.axis=1.4,at=seq(range[1],range[2],by=10),
     labels=c(range[1],rep("",length(seq(range[1],range[2],by=10))-3),100,""),
     pos=1971)
title(ylab="Relative survival (%)", line=0.5, cex.lab=1.5,adj= 0.4)
axis(side=1,las=1,cex.axis=1.4,at=c(seq(1980,2020,by=10)),pos=range[1],
     labels = c(rep("",5)))
lines(c(1971,2020),c(range[1],range[1]))

# legend
xpo=0
ypo=0.2
xx=1;yy=range[1]+scal*ypo
rect(1996+xpo,yy+0.035*scal,2017.9+xpo,yy-0.15*scal,col="white",border="grey50",lwd=0.8)
repeat{
  points(2000+xpo,yy,pch=17,col=cola[xx],cex=1.2)
  lines(c(1997+xpo,2003+xpo),c(yy,yy),lwd=1.6,col=cola[xx],lend=1)
  xx<-xx+1;yy=yy-(scal*0.058);if(xx>3){break}}
xx=1;yy=range[1]+scal*ypo
text(2010+xpo,yy,"1-year",col=cola[xx],cex=1.25);xx=xx+1;yy=yy-(scal*0.058)
text(2010+xpo,yy,"5/1-year",col=cola[xx],cex=1.25);xx=xx+1;yy=yy-(scal*0.058)
text(2010+xpo,yy,"5-year",col=cola[xx],cex=1.25);xx=xx+1;yy=yy-(scal*0.058)
text(1974,range[2]-0.05*scal,"a",cex=2.2)

#### slope plot --------------
range<-range_c;scal=scal_c
par(mar=c(2.5,2.1,0,0))

plot(NULL,xlim=c(1971,2020),ylim=c(range[1],range[2]),xlab="",ylab=""
     ,las=1, axes=FALSE)
rect(1971,range[2],2020,range[1],col="grey90",border=NA)
x<-range[1]
repeat{
  lines(c(1971,2020),c(x,x),col="white")
  x=x+1;if(x>100){break}}

x<-1980
repeat{
  lines(c(x,x),c(range[1],range[2]),col="white")
  x=x+10;if(x>2020){break}}

xx=1
polyg_slope(post_testes_fin_1y);xx=xx+1
polyg_slope(post_testes_fin_cond);xx=xx+1
polyg_slope(post_testes_fin_5y)

xx=1
slope_fit(post_testes_fin_1y);xx=xx+1
slope_fit(post_testes_fin_cond);xx=xx+1
slope_fit(post_testes_fin_5y)

axis(2,las=2,cex.axis=1.4,at=c(seq(range[1],range[2],by=1)),
     labels=c(range[1],rep("",length(seq(range[1],range[2],by=1))-2),range[2]),pos=1971)
lines(c(1971,2020),c(0,0),col="grey50")

title(ylab=expression(paste(delta, " (survival)")), line=0.5, cex.lab=1.5,adj= 0.4)

axis(side=1,las=1,cex.axis=1.4,at=c(seq(1980,2020,by=10)),pos=range[1],labels=c(rep("",5)))
text(c(seq(1978,2020,by=20)),c(rep(range[1]-0.085*scal,3)),c(seq(1980,2020,by=20)),xpd=TRUE,cex=1.4);
lines(c(1971,2020),c(range[1],range[1]))
title(xlab="Year", line=1.4, cex.lab=1.4,xpd=TRUE)


### prostate  % ------------------------------------------------------
range<-range_b;scal=scal_b
par(mar=c(0,2.1,0,0))

plot(NULL,xlim=c(1971,2020),ylim=c(range[1],range[2]),xlab="",ylab=""
     ,las=1, axes=FALSE)
rect(1971,range[2],2020,range[1],col="grey90",border=NA)
x<-10
repeat{
  lines(c(1971,2020),c(x,x),col="white")
  x=x+10;if(x>100){break}}

x<-1980
repeat{
  lines(c(x,x),c(range[1],range[2]),col="white")
  x=x+10;if(x>2020){break}}

# breaking points identification
xx=1
breakpo(post_prostate_fin_1y,-1)
xx=xx+1
breakpo(post_prostate_fin_cond,0)
xx=xx+1
breakpo(post_prostate_fin_5y,1)

# 95% credible interval for survival
xx=1
polyg_surv(post_prostate_fin_1y);xx=xx+1
polyg_surv(post_prostate_fin_cond);xx=xx+1
polyg_surv(post_prostate_fin_5y)

# data points of estimated survival
xx=1
points(prostate[prostate$shou=="fin_",]$surv_1y~prostate[prostate$shou=="fin_",]$year
       ,pch=17,col=colc[xx],cex=1);xx=xx+1
points(prostate[prostate$shou=="fin_",]$surv_cond~prostate[prostate$shou=="fin_",]$year
       ,pch=17,col=colc[xx],cex=1);xx=xx+1
points(prostate[prostate$shou=="fin_",]$surv_5y~prostate[prostate$shou=="fin_",]$year
       ,pch=17,col=colc[xx],cex=1)

# fitted lines for survival
xx=1
surv_fit(post_prostate_fin_1y);xx=xx+1
surv_fit(post_prostate_fin_cond);xx=xx+1
surv_fit(post_prostate_fin_5y);xx=xx+1

axis(2,las=2,cex.axis=1.4,at=seq(range[1],range[2],by=10),
     labels=c(range[1],rep("",length(seq(range[1],range[2],by=10))-3),100,""),
     pos=1971)
axis(side=1,las=1,cex.axis=1.4,at=c(seq(1980,2020,by=10)),pos=range[1],
     labels = c(rep("",5)))
lines(c(1971,2020),c(range[1],range[1]))
text(1974,range[2]-0.05*scal,"b",cex=2.2)

#### slope plot --------------
range<-range_c;scal=scal_c
par(mar=c(2.5,2.1,0,0))

plot(NULL,xlim=c(1971,2020),ylim=c(range[1],range[2]),xlab="",ylab=""
     ,las=1, axes=FALSE)
rect(1971,range[2],2020,range[1],col="grey90",border=NA)
x<-range[1]
repeat{
  lines(c(1971,2020),c(x,x),col="white")
  x=x+1;if(x>100){break}}

x<-1980
repeat{
  lines(c(x,x),c(range[1],range[2]),col="white")
  x=x+10;if(x>2020){break}}

xx=1
polyg_slope(post_prostate_fin_1y);xx=xx+1
polyg_slope(post_prostate_fin_cond);xx=xx+1
polyg_slope(post_prostate_fin_5y)

xx=1
slope_fit(post_prostate_fin_1y);xx=xx+1
slope_fit(post_prostate_fin_cond);xx=xx+1
slope_fit(post_prostate_fin_5y)

axis(2,las=2,cex.axis=1.4,at=c(seq(range[1],range[2],by=1)),
     labels=c(range[1],rep("",length(seq(range[1],range[2],by=1))-2),range[2]),pos=1971)
lines(c(1971,2020),c(0,0),col="grey50")

axis(side=1,las=1,cex.axis=1.4,at=c(seq(1980,2020,by=10)),pos=range[1],labels=c(rep("",5)))
text(c(seq(1978,2020,by=20)),c(rep(range[1]-0.085*scal,3)),c(seq(1980,2020,by=20)),xpd=TRUE,cex=1.4);
lines(c(1971,2020),c(range[1],range[1]))
title(xlab="Year", line=1.4, cex.lab=1.4,xpd=TRUE)


### penis  % ------------------------------------------------------
range<-range_b;scal=scal_b
par(mar=c(0,2.1,0,0))

plot(NULL,xlim=c(1971,2020),ylim=c(range[1],range[2]),xlab="",ylab=""
     ,las=1, axes=FALSE)
rect(1971,range[2],2020,range[1],col="grey90",border=NA)
x<-10
repeat{
  lines(c(1971,2020),c(x,x),col="white")
  x=x+10;if(x>100){break}}

x<-1980
repeat{
  lines(c(x,x),c(range[1],range[2]),col="white")
  x=x+10;if(x>2020){break}}

# breaking points identification
xx=1
breakpo(post_penis_fin_1y,-1)
xx=xx+1
breakpo(post_penis_fin_cond,0)
xx=xx+1
breakpo(post_penis_fin_5y,1)

# 95% credible interval for survival
xx=1
polyg_surv(post_penis_fin_1y);xx=xx+1
polyg_surv(post_penis_fin_cond);xx=xx+1
polyg_surv(post_penis_fin_5y)

# data points of estimated survival
xx=1
points(penis[penis$shou=="fin_",]$surv_1y~penis[penis$shou=="fin_",]$year
       ,pch=17,col=colc[xx],cex=1);xx=xx+1
points(penis[penis$shou=="fin_",]$surv_cond~penis[penis$shou=="fin_",]$year
       ,pch=17,col=colc[xx],cex=1);xx=xx+1
points(penis[penis$shou=="fin_",]$surv_5y~penis[penis$shou=="fin_",]$year
       ,pch=17,col=colc[xx],cex=1)

# fitted lines for survival
xx=1
surv_fit(post_penis_fin_1y);xx=xx+1
surv_fit(post_penis_fin_cond);xx=xx+1
surv_fit(post_penis_fin_5y);xx=xx+1

axis(2,las=2,cex.axis=1.4,at=seq(range[1],range[2],by=10),
     labels=c(range[1],rep("",length(seq(range[1],range[2],by=10))-3),100,""),
     pos=1971)
axis(side=1,las=1,cex.axis=1.4,at=c(seq(1980,2020,by=10)),pos=range[1],
     labels = c(rep("",5)))
lines(c(1971,2020),c(range[1],range[1]))
text(1974,range[2]-0.05*scal,"c",cex=2.2)

ypo=-60
xfl=48
rect(1956+xfl,range[1]+scal*0.965+ypo,1968+xfl,range[1]+scal*0.84+ypo,col="white",border="grey50")
rect(1956+xfl,range[1]+scal*0.9135+ypo,1968+xfl,range[1]+scal*0.8865+ypo,col="dodgerblue3",border=NA)
rect(1960+xfl,range[1]+scal*0.965+ypo,1962+xfl,range[1]+scal*0.84+ypo,col="dodgerblue3",border=NA)
#### slope plot --------------
range<-range_c;scal=scal_c
par(mar=c(2.5,2.1,0,0))

plot(NULL,xlim=c(1971,2020),ylim=c(range[1],range[2]),xlab="",ylab=""
     ,las=1, axes=FALSE)
rect(1971,range[2],2020,range[1],col="grey90",border=NA)
x<-range[1]
repeat{
  lines(c(1971,2020),c(x,x),col="white")
  x=x+1;if(x>100){break}}

x<-1980
repeat{
  lines(c(x,x),c(range[1],range[2]),col="white")
  x=x+10;if(x>2020){break}}

xx=1
polyg_slope(post_penis_fin_1y);xx=xx+1
polyg_slope(post_penis_fin_cond);xx=xx+1
polyg_slope(post_penis_fin_5y)

xx=1
slope_fit(post_penis_fin_1y);xx=xx+1
slope_fit(post_penis_fin_cond);xx=xx+1
slope_fit(post_penis_fin_5y)

axis(2,las=2,cex.axis=1.4,at=c(seq(range[1],range[2],by=1)),
     labels=c(range[1],rep("",length(seq(range[1],range[2],by=1))-2),range[2]),pos=1971)
lines(c(1971,2020),c(0,0),col="grey50")

axis(side=1,las=1,cex.axis=1.4,at=c(seq(1980,2020,by=10)),pos=range[1],labels=c(rep("",5)))
text(c(seq(1978,2020,by=20)),c(rep(range[1]-0.085*scal,3)),c(seq(1980,2020,by=20)),xpd=TRUE,cex=1.4);
lines(c(1971,2020),c(range[1],range[1]))
title(xlab="Year", line=1.4, cex.lab=1.4,xpd=TRUE)
```

### Norway

```{r, fig.height=5.75,fig.width=8.13, fig.cap="Fig. 3. Relative 1- , 5/1- (conditional) and 5-years survival in Norwish men with (a) testicular, (b) prostate and (c) penile cancer. The vertical lines mark a significant change in the survival trends (‘breaking points’) and the bottom curves show estimated annual changes in survival. The curves are solid if there is >95% plausibility of the growth or decline. Shadow areas indicate 95% credible interval. All curves are color coded (see the insert)",error=TRUE,warning=FALSE,results=FALSE,message=FALSE}
## NOR figures -----------------
m <- matrix(c(1,2,3
              ,4,6,8
              ,5,7,9), nrow = 3, ncol =3 ,byrow = TRUE)
layout(mat = m,heights = c(0.03,0.6,0.37))
par(mgp=c(1.6,0.62,0))
par(mar=c(0,0,0,0))


plot(NULL, axes=FALSE,xlab="",ylab="",xlim=c(-1,1),ylim=c(-0.85,0.85))
text(0,-0.2,"Testicular cancer",cex=1.6,font=3,xpd=TRUE)

plot(NULL, axes=FALSE,xlab="",ylab="",xlim=c(-1,1),ylim=c(-0.85,0.85))
text(0,-0.2,"Prostate cancer",cex=1.6,font=3,xpd=TRUE)

plot(NULL, axes=FALSE,xlab="",ylab="",xlim=c(-1,1),ylim=c(-0.85,0.85))
text(0,-0.2,"Penile cancer",cex=1.6,font=3,xpd=TRUE)

range_b<-c(30,110);scal_b<-range_b[2]-range_b[1]
range_c<-c(-1,4);scal_c<-range_c[2]-range_c[1]

#
### testes  % ------------------------------------------------------
range<-range_b;scal=scal_b
par(mar=c(0,2.1,0,0))

plot(NULL,xlim=c(1971,2020),ylim=c(range[1],range[2]),xlab="",ylab=""
     ,las=1, axes=FALSE)
rect(1971,range[2],2020,range[1],col="grey90",border=NA)
x<-10
repeat{
  lines(c(1971,2020),c(x,x),col="white")
  x=x+10;if(x>100){break}}

x<-1980
repeat{
  lines(c(x,x),c(range[1],range[2]),col="white")
  x=x+10;if(x>2020){break}}

# breaking points identification
xx=1
breakpo(post_testes_nor_1y,-0.8)
xx=xx+1
breakpo(post_testes_nor_cond,0.8)
xx=xx+1
breakpo(post_testes_nor_5y,0)

# 95% credible interval for survival
xx=1
polyg_surv(post_testes_nor_1y);xx=xx+1
polyg_surv(post_testes_nor_cond);xx=xx+1
polyg_surv(post_testes_nor_5y)

# data points of estimated survival
xx=1
points(testes[testes$shou=="nor_",]$surv_1y~testes[testes$shou=="nor_",]$year
       ,pch=17,col=colc[xx],cex=1);xx=xx+1
points(testes[testes$shou=="nor_",]$surv_cond~testes[testes$shou=="nor_",]$year
       ,pch=17,col=colc[xx],cex=1);xx=xx+1
points(testes[testes$shou=="nor_",]$surv_5y~testes[testes$shou=="nor_",]$year
       ,pch=17,col=colc[xx],cex=1)

# fitted lines for survival
xx=1
surv_fit(post_testes_nor_1y);xx=xx+1
surv_fit(post_testes_nor_cond);xx=xx+1
surv_fit(post_testes_nor_5y);xx=xx+1

axis(2,las=2,cex.axis=1.4,at=seq(range[1],range[2],by=10),
     labels=c(range[1],rep("",length(seq(range[1],range[2],by=10))-3),100,""),
     pos=1971)
title(ylab="Relative survival (%)", line=0.5, cex.lab=1.5,adj= 0.4)
axis(side=1,las=1,cex.axis=1.4,at=c(seq(1980,2020,by=10)),pos=range[1],
     labels = c(rep("",5)))
lines(c(1971,2020),c(range[1],range[1]))

# legend
xpo=0
ypo=0.2
xx=1;yy=range[1]+scal*ypo
rect(1996+xpo,yy+0.035*scal,2017.9+xpo,yy-0.15*scal,col="white",border="grey50",lwd=0.8)
repeat{
  points(2000+xpo,yy,pch=17,col=cola[xx],cex=1.2)
  lines(c(1997+xpo,2003+xpo),c(yy,yy),lwd=1.6,col=cola[xx],lend=1)
  xx<-xx+1;yy=yy-(scal*0.058);if(xx>3){break}}
xx=1;yy=range[1]+scal*ypo
text(2010+xpo,yy,"1-year",col=cola[xx],cex=1.25);xx=xx+1;yy=yy-(scal*0.058)
text(2010+xpo,yy,"5/1-year",col=cola[xx],cex=1.25);xx=xx+1;yy=yy-(scal*0.058)
text(2010+xpo,yy,"5-year",col=cola[xx],cex=1.25);xx=xx+1;yy=yy-(scal*0.058)
text(1974,range[2]-0.05*scal,"a",cex=2.2)

#### slope plot --------------
range<-range_c;scal=scal_c
par(mar=c(2.5,2.1,0,0))

plot(NULL,xlim=c(1971,2020),ylim=c(range[1],range[2]),xlab="",ylab=""
     ,las=1, axes=FALSE)
rect(1971,range[2],2020,range[1],col="grey90",border=NA)
x<-range[1]
repeat{
  lines(c(1971,2020),c(x,x),col="white")
  x=x+1;if(x>100){break}}

x<-1980
repeat{
  lines(c(x,x),c(range[1],range[2]),col="white")
  x=x+10;if(x>2020){break}}

xx=1
polyg_slope(post_testes_nor_1y);xx=xx+1
polyg_slope(post_testes_nor_cond);xx=xx+1
polyg_slope(post_testes_nor_5y)

xx=1
slope_fit(post_testes_nor_1y);xx=xx+1
slope_fit(post_testes_nor_cond);xx=xx+1
slope_fit(post_testes_nor_5y)

axis(2,las=2,cex.axis=1.4,at=c(seq(range[1],range[2],by=1)),
     labels=c(range[1],rep("",length(seq(range[1],range[2],by=1))-2),range[2]),pos=1971)
lines(c(1971,2020),c(0,0),col="grey50")

title(ylab=expression(paste(delta, " (survival)")), line=0.5, cex.lab=1.5,adj= 0.4)

axis(side=1,las=1,cex.axis=1.4,at=c(seq(1980,2020,by=10)),pos=range[1],labels=c(rep("",5)))
text(c(seq(1978,2020,by=20)),c(rep(range[1]-0.085*scal,3)),c(seq(1980,2020,by=20)),xpd=TRUE,cex=1.4);
lines(c(1971,2020),c(range[1],range[1]))
title(xlab="Year", line=1.4, cex.lab=1.4,xpd=TRUE)


### prostate  % ------------------------------------------------------
range<-range_b;scal=scal_b
par(mar=c(0,2.1,0,0))

plot(NULL,xlim=c(1971,2020),ylim=c(range[1],range[2]),xlab="",ylab=""
     ,las=1, axes=FALSE)
rect(1971,range[2],2020,range[1],col="grey90",border=NA)
x<-10
repeat{
  lines(c(1971,2020),c(x,x),col="white")
  x=x+10;if(x>100){break}}

x<-1980
repeat{
  lines(c(x,x),c(range[1],range[2]),col="white")
  x=x+10;if(x>2020){break}}

# breaking points identification
xx=1
breakpo(post_prostate_nor_1y,-0.8)
xx=xx+1
breakpo(post_prostate_nor_cond,-1.5)
xx=xx+1
breakpo(post_prostate_nor_5y,1.4)

# 95% credible interval for survival
xx=1
polyg_surv(post_prostate_nor_1y);xx=xx+1
polyg_surv(post_prostate_nor_cond);xx=xx+1
polyg_surv(post_prostate_nor_5y)

# data points of estimated survival
xx=1
points(prostate[prostate$shou=="nor_",]$surv_1y~prostate[prostate$shou=="nor_",]$year
       ,pch=17,col=colc[xx],cex=1);xx=xx+1
points(prostate[prostate$shou=="nor_",]$surv_cond~prostate[prostate$shou=="nor_",]$year
       ,pch=17,col=colc[xx],cex=1);xx=xx+1
points(prostate[prostate$shou=="nor_",]$surv_5y~prostate[prostate$shou=="nor_",]$year
       ,pch=17,col=colc[xx],cex=1)

# fitted lines for survival
xx=1
surv_fit(post_prostate_nor_1y);xx=xx+1
surv_fit(post_prostate_nor_cond);xx=xx+1
surv_fit(post_prostate_nor_5y);xx=xx+1

axis(2,las=2,cex.axis=1.4,at=seq(range[1],range[2],by=10),
     labels=c(range[1],rep("",length(seq(range[1],range[2],by=10))-3),100,""),
     pos=1971)
axis(side=1,las=1,cex.axis=1.4,at=c(seq(1980,2020,by=10)),pos=range[1],
     labels = c(rep("",5)))
lines(c(1971,2020),c(range[1],range[1]))
text(1974,range[2]-0.05*scal,"b",cex=2.2)

#### slope plot --------------
range<-range_c;scal=scal_c
par(mar=c(2.5,2.1,0,0))

plot(NULL,xlim=c(1971,2020),ylim=c(range[1],range[2]),xlab="",ylab=""
     ,las=1, axes=FALSE)
rect(1971,range[2],2020,range[1],col="grey90",border=NA)
x<-range[1]
repeat{
  lines(c(1971,2020),c(x,x),col="white")
  x=x+1;if(x>100){break}}

x<-1980
repeat{
  lines(c(x,x),c(range[1],range[2]),col="white")
  x=x+10;if(x>2020){break}}

xx=1
polyg_slope(post_prostate_nor_1y);xx=xx+1
polyg_slope(post_prostate_nor_cond);xx=xx+1
polyg_slope(post_prostate_nor_5y)

xx=1
slope_fit(post_prostate_nor_1y);xx=xx+1
slope_fit(post_prostate_nor_cond);xx=xx+1
slope_fit(post_prostate_nor_5y)

axis(2,las=2,cex.axis=1.4,at=c(seq(range[1],range[2],by=1)),
     labels=c(range[1],rep("",length(seq(range[1],range[2],by=1))-2),range[2]),pos=1971)
lines(c(1971,2020),c(0,0),col="grey50")

axis(side=1,las=1,cex.axis=1.4,at=c(seq(1980,2020,by=10)),pos=range[1],labels=c(rep("",5)))
text(c(seq(1978,2020,by=20)),c(rep(range[1]-0.085*scal,3)),c(seq(1980,2020,by=20)),xpd=TRUE,cex=1.4);
lines(c(1971,2020),c(range[1],range[1]))
title(xlab="Year", line=1.4, cex.lab=1.4,xpd=TRUE)


### penis  % ------------------------------------------------------
range<-range_b;scal=scal_b
par(mar=c(0,2.1,0,0))

plot(NULL,xlim=c(1971,2020),ylim=c(range[1],range[2]),xlab="",ylab=""
     ,las=1, axes=FALSE)
rect(1971,range[2],2020,range[1],col="grey90",border=NA)
x<-10
repeat{
  lines(c(1971,2020),c(x,x),col="white")
  x=x+10;if(x>100){break}}

x<-1980
repeat{
  lines(c(x,x),c(range[1],range[2]),col="white")
  x=x+10;if(x>2020){break}}

# breaking points identification
xx=1
breakpo(post_penis_nor_1y,-1)
xx=xx+1
breakpo(post_penis_nor_cond,0)
xx=xx+1
breakpo(post_penis_nor_5y,1)

# 95% credible interval for survival
xx=1
polyg_surv(post_penis_nor_1y);xx=xx+1
polyg_surv(post_penis_nor_cond);xx=xx+1
polyg_surv(post_penis_nor_5y)

# data points of estimated survival
xx=1
points(penis[penis$shou=="nor_",]$surv_1y~penis[penis$shou=="nor_",]$year
       ,pch=17,col=colc[xx],cex=1);xx=xx+1
points(penis[penis$shou=="nor_",]$surv_cond~penis[penis$shou=="nor_",]$year
       ,pch=17,col=colc[xx],cex=1);xx=xx+1
points(penis[penis$shou=="nor_",]$surv_5y~penis[penis$shou=="nor_",]$year
       ,pch=17,col=colc[xx],cex=1)

# fitted lines for survival
xx=1
surv_fit(post_penis_nor_1y);xx=xx+1
surv_fit(post_penis_nor_cond);xx=xx+1
surv_fit(post_penis_nor_5y);xx=xx+1

axis(2,las=2,cex.axis=1.4,at=seq(range[1],range[2],by=10),
     labels=c(range[1],rep("",length(seq(range[1],range[2],by=10))-3),100,""),
     pos=1971)
axis(side=1,las=1,cex.axis=1.4,at=c(seq(1980,2020,by=10)),pos=range[1],
     labels = c(rep("",5)))
lines(c(1971,2020),c(range[1],range[1]))
text(1974,range[2]-0.05*scal,"c",cex=2.2)

ypo=-60
xfl=48
rect(1956+xfl,range[1]+scal*0.965+ypo,1968+xfl,range[1]+scal*0.84+ypo,col="red2",border="grey50")
rect(1956+xfl,range[1]+scal*0.9285+ypo,1968+xfl,range[1]+scal*0.8735+ypo,col="white",border=NA)
rect(1959+xfl,range[1]+scal*0.965+ypo,1963+xfl,range[1]+scal*0.84+ypo,col="white",border=NA)
rect(1956+xfl,range[1]+scal*0.9135+ypo,1968+xfl,range[1]+scal*0.8865+ypo,col="darkblue",border=NA)
rect(1960+xfl,range[1]+scal*0.965+ypo,1962+xfl,range[1]+scal*0.84+ypo,col="darkblue",border=NA)

#### slope plot --------------
range<-range_c;scal=scal_c
par(mar=c(2.5,2.1,0,0))

plot(NULL,xlim=c(1971,2020),ylim=c(range[1],range[2]),xlab="",ylab=""
     ,las=1, axes=FALSE)
rect(1971,range[2],2020,range[1],col="grey90",border=NA)
x<-range[1]
repeat{
  lines(c(1971,2020),c(x,x),col="white")
  x=x+1;if(x>100){break}}

x<-1980
repeat{
  lines(c(x,x),c(range[1],range[2]),col="white")
  x=x+10;if(x>2020){break}}

xx=1
polyg_slope(post_penis_nor_1y);xx=xx+1
polyg_slope(post_penis_nor_cond);xx=xx+1
polyg_slope(post_penis_nor_5y)

xx=1
slope_fit(post_penis_nor_1y);xx=xx+1
slope_fit(post_penis_nor_cond);xx=xx+1
slope_fit(post_penis_nor_5y)

axis(2,las=2,cex.axis=1.4,at=c(seq(range[1],range[2],by=1)),
     labels=c(range[1],rep("",length(seq(range[1],range[2],by=1))-2),range[2]),pos=1971)
lines(c(1971,2020),c(0,0),col="grey50")

axis(side=1,las=1,cex.axis=1.4,at=c(seq(1980,2020,by=10)),pos=range[1],labels=c(rep("",5)))
text(c(seq(1978,2020,by=20)),c(rep(range[1]-0.085*scal,3)),c(seq(1980,2020,by=20)),xpd=TRUE,cex=1.4);
lines(c(1971,2020),c(range[1],range[1]))
title(xlab="Year", line=1.4, cex.lab=1.4,xpd=TRUE)
```

### Sweden

```{r, fig.height=5.75,fig.width=8.13, fig.cap="Fig. 4. Relative 1- , 5/1- (conditional) and 5-years survival in Swedish men with (a) testicular, (b) prostate and (c) penile cancer. The vertical lines mark a significant change in the survival trends (‘breaking points’) and the bottom curves show estimated annual changes in survival. The curves are solid if there is >95% plausibility of the growth or decline. Shadow areas indicate 95% credible interval. All curves are color coded (see the insert)",error=TRUE,warning=FALSE,results=FALSE,message=FALSE}
## SWE figures -----------------
m <- matrix(c(1,2,3
              ,4,6,8
              ,5,7,9), nrow = 3, ncol =3 ,byrow = TRUE)
layout(mat = m,heights = c(0.03,0.6,0.37))
par(mgp=c(1.6,0.62,0))
par(mar=c(0,0,0,0))


plot(NULL, axes=FALSE,xlab="",ylab="",xlim=c(-1,1),ylim=c(-0.85,0.85))
text(0,-0.2,"Testicular cancer",cex=1.6,font=3,xpd=TRUE)

plot(NULL, axes=FALSE,xlab="",ylab="",xlim=c(-1,1),ylim=c(-0.85,0.85))
text(0,-0.2,"Prostate cancer",cex=1.6,font=3,xpd=TRUE)

plot(NULL, axes=FALSE,xlab="",ylab="",xlim=c(-1,1),ylim=c(-0.85,0.85))
text(0,-0.2,"Penile cancer",cex=1.6,font=3,xpd=TRUE)

range_b<-c(30,110);scal_b<-range_b[2]-range_b[1]
range_c<-c(-1,4);scal_c<-range_c[2]-range_c[1]

#
### testes  % ------------------------------------------------------
range<-range_b;scal=scal_b
par(mar=c(0,2.1,0,0))

plot(NULL,xlim=c(1971,2020),ylim=c(range[1],range[2]),xlab="",ylab=""
     ,las=1, axes=FALSE)
rect(1971,range[2],2020,range[1],col="grey90",border=NA)
x<-10
repeat{
  lines(c(1971,2020),c(x,x),col="white")
  x=x+10;if(x>100){break}}

x<-1980
repeat{
  lines(c(x,x),c(range[1],range[2]),col="white")
  x=x+10;if(x>2020){break}}

# breaking points identification
xx=1
breakpo(post_testes_swe_1y,-0.6)
xx=xx+1
breakpo(post_testes_swe_cond,0)
xx=xx+1
breakpo(post_testes_swe_5y,0.6)

# 95% credible interval for survival
xx=1
polyg_surv(post_testes_swe_1y);xx=xx+1
polyg_surv(post_testes_swe_cond);xx=xx+1
polyg_surv(post_testes_swe_5y)

# data points of estimated survival
xx=1
points(testes[testes$shou=="swe_",]$surv_1y~testes[testes$shou=="swe_",]$year
       ,pch=17,col=colc[xx],cex=1);xx=xx+1
points(testes[testes$shou=="swe_",]$surv_cond~testes[testes$shou=="swe_",]$year
       ,pch=17,col=colc[xx],cex=1);xx=xx+1
points(testes[testes$shou=="swe_",]$surv_5y~testes[testes$shou=="swe_",]$year
       ,pch=17,col=colc[xx],cex=1)

# fitted lines for survival
xx=1
surv_fit(post_testes_swe_1y);xx=xx+1
surv_fit(post_testes_swe_cond);xx=xx+1
surv_fit(post_testes_swe_5y);xx=xx+1

axis(2,las=2,cex.axis=1.4,at=seq(range[1],range[2],by=10),
     labels=c(range[1],rep("",length(seq(range[1],range[2],by=10))-3),100,""),
     pos=1971)
title(ylab="Relative survival (%)", line=0.5, cex.lab=1.5,adj= 0.4)
axis(side=1,las=1,cex.axis=1.4,at=c(seq(1980,2020,by=10)),pos=range[1],
     labels = c(rep("",5)))
lines(c(1971,2020),c(range[1],range[1]))

# legend
xpo=0
ypo=0.2
xx=1;yy=range[1]+scal*ypo
rect(1996+xpo,yy+0.035*scal,2017.9+xpo,yy-0.15*scal,col="white",border="grey50",lwd=0.8)
repeat{
  points(2000+xpo,yy,pch=17,col=cola[xx],cex=1.2)
  lines(c(1997+xpo,2003+xpo),c(yy,yy),lwd=1.6,col=cola[xx],lend=1)
  xx<-xx+1;yy=yy-(scal*0.058);if(xx>3){break}}
xx=1;yy=range[1]+scal*ypo
text(2010+xpo,yy,"1-year",col=cola[xx],cex=1.25);xx=xx+1;yy=yy-(scal*0.058)
text(2010+xpo,yy,"5/1-year",col=cola[xx],cex=1.25);xx=xx+1;yy=yy-(scal*0.058)
text(2010+xpo,yy,"5-year",col=cola[xx],cex=1.25);xx=xx+1;yy=yy-(scal*0.058)
text(1974,range[2]-0.05*scal,"a",cex=2.2)

#### slope plot --------------
range<-range_c;scal=scal_c
par(mar=c(2.5,2.1,0,0))

plot(NULL,xlim=c(1971,2020),ylim=c(range[1],range[2]),xlab="",ylab=""
     ,las=1, axes=FALSE)
rect(1971,range[2],2020,range[1],col="grey90",border=NA)
x<-range[1]
repeat{
  lines(c(1971,2020),c(x,x),col="white")
  x=x+1;if(x>100){break}}

x<-1980
repeat{
  lines(c(x,x),c(range[1],range[2]),col="white")
  x=x+10;if(x>2020){break}}

xx=1
polyg_slope(post_testes_swe_1y);xx=xx+1
polyg_slope(post_testes_swe_cond);xx=xx+1
polyg_slope(post_testes_swe_5y)

xx=1
slope_fit(post_testes_swe_1y);xx=xx+1
slope_fit(post_testes_swe_cond);xx=xx+1
slope_fit(post_testes_swe_5y)

axis(2,las=2,cex.axis=1.4,at=c(seq(range[1],range[2],by=1)),
     labels=c(range[1],rep("",length(seq(range[1],range[2],by=1))-2),range[2]),pos=1971)
lines(c(1971,2020),c(0,0),col="grey50")

title(ylab=expression(paste(delta, " (survival)")), line=0.5, cex.lab=1.5,adj= 0.4)

axis(side=1,las=1,cex.axis=1.4,at=c(seq(1980,2020,by=10)),pos=range[1],labels=c(rep("",5)))
text(c(seq(1978,2020,by=20)),c(rep(range[1]-0.085*scal,3)),c(seq(1980,2020,by=20)),xpd=TRUE,cex=1.4);
lines(c(1971,2020),c(range[1],range[1]))
title(xlab="Year", line=1.4, cex.lab=1.4,xpd=TRUE)


### prostate  % ------------------------------------------------------
range<-range_b;scal=scal_b
par(mar=c(0,2.1,0,0))

plot(NULL,xlim=c(1971,2020),ylim=c(range[1],range[2]),xlab="",ylab=""
     ,las=1, axes=FALSE)
rect(1971,range[2],2020,range[1],col="grey90",border=NA)
x<-10
repeat{
  lines(c(1971,2020),c(x,x),col="white")
  x=x+10;if(x>100){break}}

x<-1980
repeat{
  lines(c(x,x),c(range[1],range[2]),col="white")
  x=x+10;if(x>2020){break}}

# breaking points identification
xx=1
breakpo(post_prostate_swe_1y,-1.4)
xx=xx+1
breakpo(post_prostate_swe_cond,-0.2)
xx=xx+1
breakpo(post_prostate_swe_5y,0.8)

# 95% credible interval for survival
xx=1
polyg_surv(post_prostate_swe_1y);xx=xx+1
polyg_surv(post_prostate_swe_cond);xx=xx+1
polyg_surv(post_prostate_swe_5y)

# data points of estimated survival
xx=1
points(prostate[prostate$shou=="swe_",]$surv_1y~prostate[prostate$shou=="swe_",]$year
       ,pch=17,col=colc[xx],cex=1);xx=xx+1
points(prostate[prostate$shou=="swe_",]$surv_cond~prostate[prostate$shou=="swe_",]$year
       ,pch=17,col=colc[xx],cex=1);xx=xx+1
points(prostate[prostate$shou=="swe_",]$surv_5y~prostate[prostate$shou=="swe_",]$year
       ,pch=17,col=colc[xx],cex=1)

# fitted lines for survival
xx=1
surv_fit(post_prostate_swe_1y);xx=xx+1
surv_fit(post_prostate_swe_cond);xx=xx+1
surv_fit(post_prostate_swe_5y);xx=xx+1

axis(2,las=2,cex.axis=1.4,at=seq(range[1],range[2],by=10),
     labels=c(range[1],rep("",length(seq(range[1],range[2],by=10))-3),100,""),
     pos=1971)
axis(side=1,las=1,cex.axis=1.4,at=c(seq(1980,2020,by=10)),pos=range[1],
     labels = c(rep("",5)))
lines(c(1971,2020),c(range[1],range[1]))
text(1974,range[2]-0.05*scal,"b",cex=2.2)

#### slope plot --------------
range<-range_c;scal=scal_c
par(mar=c(2.5,2.1,0,0))

plot(NULL,xlim=c(1971,2020),ylim=c(range[1],range[2]),xlab="",ylab=""
     ,las=1, axes=FALSE)
rect(1971,range[2],2020,range[1],col="grey90",border=NA)
x<-range[1]
repeat{
  lines(c(1971,2020),c(x,x),col="white")
  x=x+1;if(x>100){break}}

x<-1980
repeat{
  lines(c(x,x),c(range[1],range[2]),col="white")
  x=x+10;if(x>2020){break}}

xx=1
polyg_slope(post_prostate_swe_1y);xx=xx+1
polyg_slope(post_prostate_swe_cond);xx=xx+1
polyg_slope(post_prostate_swe_5y)

xx=1
slope_fit(post_prostate_swe_1y);xx=xx+1
slope_fit(post_prostate_swe_cond);xx=xx+1
slope_fit(post_prostate_swe_5y)

axis(2,las=2,cex.axis=1.4,at=c(seq(range[1],range[2],by=1)),
     labels=c(range[1],rep("",length(seq(range[1],range[2],by=1))-2),range[2]),pos=1971)
lines(c(1971,2020),c(0,0),col="grey50")

axis(side=1,las=1,cex.axis=1.4,at=c(seq(1980,2020,by=10)),pos=range[1],labels=c(rep("",5)))
text(c(seq(1978,2020,by=20)),c(rep(range[1]-0.085*scal,3)),c(seq(1980,2020,by=20)),xpd=TRUE,cex=1.4);
lines(c(1971,2020),c(range[1],range[1]))
title(xlab="Year", line=1.4, cex.lab=1.4,xpd=TRUE)










### penis  % ------------------------------------------------------
range<-range_b;scal=scal_b
par(mar=c(0,2.1,0,0))

plot(NULL,xlim=c(1971,2020),ylim=c(range[1],range[2]),xlab="",ylab=""
     ,las=1, axes=FALSE)
rect(1971,range[2],2020,range[1],col="grey90",border=NA)
x<-10
repeat{
  lines(c(1971,2020),c(x,x),col="white")
  x=x+10;if(x>100){break}}

x<-1980
repeat{
  lines(c(x,x),c(range[1],range[2]),col="white")
  x=x+10;if(x>2020){break}}

# breaking points identification
xx=1
breakpo(post_penis_swe_1y,-1)
xx=xx+1
breakpo(post_penis_swe_cond,0)
xx=xx+1
breakpo(post_penis_swe_5y,1)

# 95% credible interval for survival
xx=1
polyg_surv(post_penis_swe_1y);xx=xx+1
polyg_surv(post_penis_swe_cond);xx=xx+1
polyg_surv(post_penis_swe_5y)

# data points of estimated survival
xx=1
points(penis[penis$shou=="swe_",]$surv_1y~penis[penis$shou=="swe_",]$year
       ,pch=17,col=colc[xx],cex=1);xx=xx+1
points(penis[penis$shou=="swe_",]$surv_cond~penis[penis$shou=="swe_",]$year
       ,pch=17,col=colc[xx],cex=1);xx=xx+1
points(penis[penis$shou=="swe_",]$surv_5y~penis[penis$shou=="swe_",]$year
       ,pch=17,col=colc[xx],cex=1)

# fitted lines for survival
xx=1
surv_fit(post_penis_swe_1y);xx=xx+1
surv_fit(post_penis_swe_cond);xx=xx+1
surv_fit(post_penis_swe_5y);xx=xx+1

axis(2,las=2,cex.axis=1.4,at=seq(range[1],range[2],by=10),
     labels=c(range[1],rep("",length(seq(range[1],range[2],by=10))-3),100,""),
     pos=1971)
axis(side=1,las=1,cex.axis=1.4,at=c(seq(1980,2020,by=10)),pos=range[1],
     labels = c(rep("",5)))
lines(c(1971,2020),c(range[1],range[1]))
text(1974,range[2]-0.05*scal,"c",cex=2.2)

ypo=-60
xfl=48
rect(1956+xfl,range[1]+scal*0.965+ypo,1968+xfl,range[1]+scal*0.84+ypo,col="blue",border="grey50")
rect(1956+xfl,range[1]+scal*0.9135+ypo,1968+xfl,range[1]+scal*0.8865+ypo,col="yellow",border=NA)
rect(1960+xfl,range[1]+scal*0.965+ypo,1962+xfl,range[1]+scal*0.84+ypo,col="yellow",border=NA)
#### slope plot --------------
range<-range_c;scal=scal_c
par(mar=c(2.5,2.1,0,0))

plot(NULL,xlim=c(1971,2020),ylim=c(range[1],range[2]),xlab="",ylab=""
     ,las=1, axes=FALSE)
rect(1971,range[2],2020,range[1],col="grey90",border=NA)
x<-range[1]
repeat{
  lines(c(1971,2020),c(x,x),col="white")
  x=x+1;if(x>100){break}}

x<-1980
repeat{
  lines(c(x,x),c(range[1],range[2]),col="white")
  x=x+10;if(x>2020){break}}

xx=1
polyg_slope(post_penis_swe_1y);xx=xx+1
polyg_slope(post_penis_swe_cond);xx=xx+1
polyg_slope(post_penis_swe_5y)

xx=1
slope_fit(post_penis_swe_1y);xx=xx+1
slope_fit(post_penis_swe_cond);xx=xx+1
slope_fit(post_penis_swe_5y)

axis(2,las=2,cex.axis=1.4,at=c(seq(range[1],range[2],by=1)),
     labels=c(range[1],rep("",length(seq(range[1],range[2],by=1))-2),range[2]),pos=1971)
lines(c(1971,2020),c(0,0),col="grey50")

axis(side=1,las=1,cex.axis=1.4,at=c(seq(1980,2020,by=10)),pos=range[1],labels=c(rep("",5)))
text(c(seq(1978,2020,by=20)),c(rep(range[1]-0.085*scal,3)),c(seq(1980,2020,by=20)),xpd=TRUE,cex=1.4);
lines(c(1971,2020),c(range[1],range[1]))
title(xlab="Year", line=1.4, cex.lab=1.4,xpd=TRUE)
```


# Reproducibility ticket
```{r error, warning=FALSE,results=TRUE,message=FALSE}
  sessioninfo::session_info()
```
  